### 2021-09-07 instructionsetsforstrangers/instructionsetsforstrangers_partI.tex
pretty lame, but it's all there

\documentclass[11pt]{report}
\usepackage{geometry}
\geometry{letterpaper, top=1in, bottom=1in, left=1in, right=1in}                                 
\usepackage[parfill]{parskip}   
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{array}
\usepackage{tabularx}
\setlength{\parindent}{0em}
\pagenumbering{gobble}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\newday}[1]{\newpage{\LARGE \textbf{#1}}\\}
\newcommand{\imagefolder}{/Users/lauriaclarke/Documents/mfadt/ms1/lauriaclarke.github.io/instructionsetsforstrangers}

\title{INSTRUCTION SETS FOR STRANGERS -- PART I}
\date{}

\begin{document}
\maketitle
%\tableofcontents
%----------------------------------------------
\section*{Location Choice}

After some discussion, we decided that a crosswalk would be the most exciting place to observe interactions between strangers. Given the high degree of randomness in who you encounter while crossing the street it seemed like an interesting choice -- particularly if the goal is to foster interaction between strangers.

To pick a crosswalk we walked along 17th street between 5th and 2nd avenues.

On the northeast corner of Union Square Park we discovered a crosswalk that included a median area. Few crosswalks in Manhattan cross two-way traffic -- this crosswalk crosses Park Ave. which is two lanes of divided and opposing traffic. Overall, this has the effect of making the intersection much larger and the act of walking across the street is much longer.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.10, angle=270]{"images/crosswalk_nopeople.jpg"}
\caption{The selected crosswalk on 17th st.}
\label{fig:nopeople}
\end{figure}

% insert history about Union Sq Park here

\section*{Initial Observation}

After walking by the first time around 4 PM on a weekday, we decided to return on a Saturday at noon -- during the Farmer's Market. The area surrounding Union Square is heavily trafficked on weekdays by commuters going to or from the train. Our hope for the weekend, was that more people would be paying attention to their surroundings and less focused on getting where they needed to go quickly. After completing a the AEIOU exercise at the intersection, this generally seemed to be true. Most people crossing the street were shopping at the market and spending time with friends.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.8]{"images/ms1_aeiou.png"}\\

\includegraphics[scale=0.8]{"images/ms1_behaviormap.png"}
\caption{AEIOU and Behavior mapping exercise}
\label{fig:behaviormap}
\end{figure}

Behavior mapping at the crosswalk revealed some more interesting data. People waiting at the crosswalk were generally found to be doing one of three things: chatting with a friend, looking at their phone, or looking at traffic. The majority of people waiting at the intersection spent their time looking at traffic. I think this may be due to the fact that this intersection is quite long -- it has a median -- and cars moving down Park Ave. tend to be going pretty fast. It seems like an intersection that requires attention as a pedestrian.

Since most of the people crossing the street here were travelling to or from the farmer's market and many were with fiends or family, the majority of foot traffic was routed over the small delta in the bike lane (see shaded shape in map) with people clustered in groups. In comparison, few people stopped on the medians in Park Ave. Those that did were generally alone and walking out of sync with the lights and were paying attention to traffic, not their cellphones.

\section*{Ideation}

After collecting data, we began to brainstorm about which interactions would be most interesting between strangers crossing the street. The idea of playing a simple game with a strange on the other side of the crosswalk was very appealing, but most required more synchronization and direction than was possible in such a short amount of time (the amount of time spent waiting at the crosswalk).

\begin{figure}[ht]
\centering
\includegraphics[width=1.4\textwidth, angle=90]{"images/ms1_brainstorming.png"}
\caption{Results of brainstorming activity.}
\label{fig:brainstorm}
\end{figure}

After deciding that the ideal interaction would be really quick -- such as a greeting -- we began thinking about fun ways to wave at someone on the other side of the street. The idea of having a giant arm and hand on each side of the street was very appealing and seemed both simple and effective. The below sketch illustrates what a pair of hands / arms would look like on location.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.08]{"images/wavemachine_overlay_2.jpg"}
\caption{Sketch of device on location.}
\label{fig:overlay}
\end{figure}

The best part of such a contraption is that it would not necessarily require the use of electronics. Using thin plywood, springs and some clever balancing, we could create the arm mechanism from very basic materials. Below is a sketch of one possible mechanism.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{"images/wavemachine_sketch.png"}
\caption{Possible mechanical design of device.}
\label{fig:sketch}
\end{figure}

\end{document}

### 2021-09-13 instructionsetsforstrangers/instructionsetsforstrangers.tex
part II

--- 2021-09-07 instructionsetsforstrangers/instructionsetsforstrangers_partI.tex
@@
 \usepackage{tabularx}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,      
+    urlcolor=cyan
+    }
+
 \setlength{\parindent}{0em}
@@
 
-\title{INSTRUCTION SETS FOR STRANGERS -- PART I}
+\title{INSTRUCTION SETS FOR STRANGERS}
 \date{}
+\author{Lauria Clarke}
 
@@
 %----------------------------------------------
-\section*{Location Choice}
+\section*{PART I - Ideation}
+\subsection*{Location Choice}
 
@@
 \centering
-\includegraphics[scale=0.10, angle=270]{"images/crosswalk_nopeople.jpg"}
+\includegraphics[scale=0.10, angle=270]{"images/I/crosswalk_nopeople.jpg"}
 \caption{The selected crosswalk on 17th st.}
@@
 
-\section*{Initial Observation}
+\subsection*{Initial Observation}
 
@@
 Our hope for the weekend, was that more people would be paying attention to their surroundings and less focused on getting where they needed to go quickly.
-After completing a the AEIOU exercise at the intersection, this generally seemed to be true.
+After completing the AEIOU exercise at the intersection, this generally seemed to be true.
 Most people crossing the street were shopping at the market and spending time with friends.
@@
 \centering
-\includegraphics[scale=0.8]{"images/ms1_aeiou.png"}\\
-
-\includegraphics[scale=0.8]{"images/ms1_behaviormap.png"}
+\includegraphics[scale=0.8]{"images/I/ms1_aeiou.png"}\\
+\vspace{0.5cm}
+\includegraphics[scale=0.8]{"images/I/ms1_behaviormap.png"}
 \caption{AEIOU and Behavior mapping exercise}
@@
 
-\section*{Ideation}
+\subsection*{Ideation}
 
@@
 \centering
-\includegraphics[width=1.4\textwidth, angle=90]{"images/ms1_brainstorming.png"}
+\includegraphics[width=1.4\textwidth, angle=90]{"images/I/ms1_brainstorming.png"}
 \caption{Results of brainstorming activity.}
@@
 \centering
-\includegraphics[scale=0.08]{"images/wavemachine_overlay_2.jpg"}
+\includegraphics[width=0.8\textwidth]{"images/I/wavemachine_overlay_2.jpg"}
 \caption{Sketch of device on location.}
@@
 \centering
-\includegraphics[scale=0.6]{"images/wavemachine_sketch.png"}
+\includegraphics[width=0.8\textwidth]{"images/I/wavemachine_sketch.png"}
 \caption{Possible mechanical design of device.}
@@
 
+\clearpage
+\section*{PART II - Iteration}
+\subsection*{prototyping}
+
+The initial prototype was extremely quick.
+It successfully proved that the basic mechanism did, in fact work.
+\href{https://drive.google.com/file/d/1A-JF5d3mOUxOp_nL0GExQ2gzeuzpgY3x/view?usp=sharing}{Link to video of initial prototype.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth, angle=-90]{"images/II/prototype_1.JPG"}
+\caption{Initial prototype.}
+\end{figure}
+
+The initial prototype required improvement in four areas; a more solid connection between the arm and the base, a way for the base to be connected to a street post securely, improved elastic, and an eye catching handle.
+
+Using a bolt and some washers the arm and the base were connected in a way that allowed the arm to swing freely without twisting.
+Next, in order to mount the wave machine, we found some large zip ties and drilled holes in the base to thread the zip ties through.
+During this process we discovered that making the mechanism one-sided decreased the complexity of the device and made it easier to mount the base.
+By restricting the movement to be one-sided we also decreased the chance of the arm becoming stuck in either direction.
+Despite the reduction in range, it was still quite satisfying to use the mechanism.
+As we played with the movement of the arm, the easiest elastics on hand were hair-ties.
+By trying a couple different sizes we honed in on the best tension for the weight of the arm.
+Finally, we cut the end off of an aluminum mop and colored it bright red with a sharpie in the hope that this would grab someone's attention.
+\href{https://drive.google.com/file/d/10h-Otp0xxu8sbUBII9ftw7f-XZoaok2H/view?usp=sharing}{Link to video of second prototype.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth, angle=-90]{"images/II/prototype_2.JPG"}
+\caption{Initial prototype.}
+\end{figure}
+
+\subsection*{testing results}
+
+Due to time constraints, we chose to test the wave machine close to home instead of at the proposed site.
+Installation went well; it was easy to attach the base securely to the street post and then adjust it to be out of the way of the crosswalk signals.
+Pulling the handle was satisfying and the motion of the mechanism was smooth and effective.
+It was fairly visible from across the street when the arm was waving.
+\href{https://drive.google.com/file/d/1Gzw53p0xpSqZ-RktlmBlWZ7QyR_WPB5E/view?usp=sharing}{Link to video of installed protptype.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.5\textwidth, angle=-90]{"images/II/install_closeup.JPG"}
+\includegraphics[width=0.5\textwidth, angle=-90]{"images/II/install_closeup_2.JPG"}
+\caption{Installed mechanism.}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth, angle=-90]{"images/II/install_far.JPG"}
+\caption{Installed mechanism.}
+\end{figure}
+
+Unfortunately, while the chosen time and location yielded a high degree of foot traffic, few noticed the wave machine.
+A few people looked at it, but no one pulled the handle.
+One person waved back at us while we were using the machine and a few others smiled.
+Other than that, the initial testing was successful for mechanical refinement, but not so much for user feedback.
+
+\subsection*{next steps}
+
+It was clear after initial testing that the wave machine needed to be more noticeable and the chosen location needed to be more heavily trafficked by individuals out looking for fun.
+In addition to making the mechanism robust enough to stand up to real use, three modifications were decided upon: improved location, more visible arm, more alluring handle.
+
+We happened to witness a particular crosswalk that was very crowded on both sides late at night.
+Nearly all of the people on both sides of the street were going to or from bars and were already pretty intoxicated.
+Many people near the crosswalk were waiting in line to enter various bars.
+Other than trying to get to the line, none of these people were in a particular rush.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.3\textwidth]{"images/II/elwire.jpg"}
+\caption{Electroluminescent wire.}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth]{"images/II/essex_magicianbar.JPG"}
+\\
+\includegraphics[width=0.8\textwidth]{"images/II/essex_opposite.JPG"}
+\caption{Nightlife on both sides of proposed crosswalk.}
+\end{figure}
+
+Using electroluminescent (EL) wire we planned to increase the visibility of the arm.
+The EL wire would be placed around the outline of the arm and hand and could be set to be constantly on, or to blink.
+In addition to grabbing peoples attention, this would create an interesting visual effect as the arm waved through the air.
+
+Lastly, the handle needed to be improved upon. To make it appealing at night we though it would be a good idea to have it light up and blink.
+
+\clearpage
+\section*{PART III - Installation}
+
 \end{document}

### 2021-09-26 tooltracker/data.js
added tool tracker

@@ -1,0 +1,71 @@
+// 0 =  snips
+// 1 =  hammer
+// 2 =  mallet
+// 3 =  drill
+// 4 =  4 mm allen
+// 5 =  5 mm allen
+// 6 =  tape measure
+// 7 =  screwdriver
+// 8 =  multimeter 
+// 9 =  staplegun
+// 10 = lights
+// 11 = mini gluegun
+// 12 = combosquare
+// 13 = level
+// 14 = clamps    
+// 15 = camera
+// 16 = xacto
+// 17 = mini hack saw
+// 18 = large hack saw
+
+var toolNames = ["snips",
+                 "hammer",
+                 "mallet",
+                 "drill",
+                 "4 mm allen",
+                 "5 mm allen",
+                 "tape measure",
+                 "screwdriver",
+                 "multimeter",
+                 "staplegun",
+                 "light",
+                 "mini gluegun",
+                 "combosquare",
+                 "level",
+                 "clamps",
+                 "camera",
+                 "xacto",
+                 "mini hack saw",   
+                 "large hack saw"];   
+
+
+var toolUse = [];
+// move in
+toolUse[0] = [1, 6];
+toolUse[1] = [1, 6, 7];
+toolUse[2] = [1, 6];
+toolUse[3] = [1, 7];
+toolUse[4] = [1, 6, 7, 5];
+// pegboard install
+toolUse[18] = [3, 13, 7];
+toolUse[19] = [3, 13, 7];
+toolUse[20] = [1];
+toolUse[21] = [1];
+// day of postcards due
+toolUse[37] = [10, 15, 16, 12];
+// first prototype
+toolUse[44] = [3, 1];
+// second prototype
+toolUse[48] = [18, 1, 3, 0, 14, 12];
+// blinds installation
+toolUse[49] = [3, 13, 7, 1, 6];
+// final prototype p1
+toolUse[50] = [3, 14, 7];
+// final prototype p2
+toolUse[51] = [3, 14, 7, 11, 0];
+toolUse[52] = [3, 13, 7];
+toolUse[53] = [3, 13, 7];
+// bike fix
+toolUse[55] = [4, 5];
+
+

### 2021-09-26 tooltracker/sketch.js

@@ -1,0 +1,35 @@
+// canvas
+var canvasX = 4000;
+var canvasY = 600;
+var nDays = 55;
+
+function setup(){
+  createCanvas(canvasX, canvasY);
+  textAlign(CENTER, CENTER);
+}
+
+function draw() {
+  var x1 = 0 + 100;
+  var y1 = canvasY / 2;
+  var x2 = 0 + 100;
+  var y2 = canvasY / 2;
+
+  increment  = (canvasX - 100 )/ (nDays + 1); 
+  console.log(toolUse.length);
+  for(let i = 0; i < toolUse.length; i++){
+    line(x1, y1, x2, y2);
+    ellipse(x2, y2, 5, 5);
+    printTools(x2, i);
+    x2 = x1
+    x1 += increment;
+  }
+  background();
+}
+
+function printTools(x, index){
+  if(toolUse[index] != undefined){
+    for(let i = 0; i < toolUse[index].length; i++){
+       text(toolNames[toolUse[index][i]], x, canvasY / 2 - (i + 1) * 20);
+    }
+  }
+}

### 2021-09-29 instructionsetsforstrangers/instructionsetsforstrangers.tex
updated index
some instruatcion sets updates

@@
 \title{INSTRUCTION SETS FOR STRANGERS}
-\date{}
+\date{\today}
 \author{Lauria Clarke}
@@
 \section*{PART II - Iteration}
-\subsection*{prototyping}
+\subsection*{initial prototypes}
 
@@
 \clearpage
+\subsection*{final prototype}
+
+Combinging the feedback from class critique we needed to move forward on the three main areas of improvemnet:
+\begin{itemize}
+\item
+  make the entire mechanism more robust for installation
+\item
+  make the hand and activation mechnaism more visible at night
+\item
+  change the activation mechnaism so that it is easier to understand how to activate the hand
+\end{itemize}
+
+First we decided to increase the scale of the wave machine and make it more robust.
+After a quick survey of the scrap bin in the wood shop we found all the materials that we needed.
+The first step was making the hands themselves.
+These were cut from two sheets of $1/16$ inch plywood on the bandsaw.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.5\textwidth, angle=270]{"images/III/plainhands.jpg"}
+\caption{Big hands!}
+\end{figure}
+
+Next we found some $1/8$ inch and $3/4$ inch plywood and created the base of the new mechanishm.
+At this point we also determined the best way to make the handle to activate the hand waving easier to use.
+A handle that needs to be pulled is ambiguous and not very visual, so we decided to make the handle another lever in the shape of a hand that would encourage users to use their hands to activate it.
+This was a simple lever that sticks out of one side of the mechanism pulled one side of the hand waving lever down when activated.
+\href{https://drive.google.com/file/d/1A4jMRefTUcI-PZPhbS6XSqFU6V4kzS8j/view?usp=sharing}{Link to video of lever action.}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.5\textwidth, angle=270]{"images/III/leverwithbase.jpg"}
+\caption{Big hands!}
+\end{figure}
+
+\clearpage
 \section*{PART III - Installation}

### 2021-09-30 instructionsetsforstrangers/instructionsetsforstrangers.tex
link to wavemachine video

@@
 At this point we also determined the best way to make the handle to activate the hand waving easier to use.
-A handle that needs to be pulled is ambiguous and not very visual, so we decided to make the handle another lever in the shape of a hand that would encourage users to use their hands to activate it.
+A handle that needs to be pulled is a bit ambiguous and not very visual, so we decided to make the handle another lever in the shape of a hand that would encourage users to use their hands to activate it.
 This was a simple lever that sticks out of one side of the mechanism pulled one side of the hand waving lever down when activated.
@@
 \includegraphics[width=0.5\textwidth, angle=270]{"images/III/leverwithbase.jpg"}
-\caption{Big hands!}
+\caption{Lever action.}
 \end{figure}
@@
 
+takeaways
+
+no one wants to push hard on the hands
+
+people who were more intoxicaed we re less gentle
+
 \end{document}

### 2021-11-02 package/sketch.js
apd push

@@ -12,6 +12,7 @@ function setup() {
   const paper_canvas = document.createElement("canvas");
   paper_canvas.setAttribute("width", "800");
   paper_canvas.setAttribute("height", "800");
+  paper_canvas.style = "position:absolute; left: 25%;";
   document.body.append(paper_canvas);
   paper.setup(paper_canvas);
 }

### 2021-11-02 package/label.js
apd push

@@ -89,7 +89,7 @@ Label.prototype.makeLabel = function(){
   }else{
     this.makeBasket();
   }
-  // this.placeText();
+  this.placeText();
   this.g.addChild(this.text);
   this.g.addChild(this.berries);
   this.g.addChild(this.basket);

### 2021-11-02 package/label.js
package

@@ -436,13 +436,13 @@ Label.prototype.placeText = function(){
     // textC.bounds.bottomCenter = this.bounds.bottomCenter;
     // textC.fontSize = fontScale * 0.07;
 
-    const textD = new paper.PointText();
-    textD.fontFamily = 'Tienne';
-    textD.fillColor = fillColor;
-    textD.content = this.nPieces + ' pieces';
-    textD.justification = 'center';
-    textD.bounds.bottomCenter = this.bounds.bottomCenter;
-    textD.fontSize = this.height / 2 * fontScale;
+   //  const textD = new paper.PointText();
+   //  textD.fontFamily = 'Tienne';
+   //  textD.fillColor = fillColor;
+   //  textD.content = this.nPieces + ' pieces';
+   //  textD.justification = 'center';
+   //  textD.bounds.bottomCenter = this.bounds.bottomCenter;
+   //  textD.fontSize = this.height / 2 * fontScale;
     // this.text.addChild(textA);
     // this.text.addChild(textB);
     // this.text.addChild(textC);
@@ -452,10 +452,10 @@ Label.prototype.placeText = function(){
     text.content = frontText;
     text.justification = 'center';
     text.fillColor = 'black';
-    text.fontSize =  this.bounds.size.width / 2 * fontScale;
-    if(text.bounds.size.width > this.bounds.size.width){
-      text.fontSize =  this.bounds.size.width / 3 * fontScale;
-    }
+    text.fontSize = 10; //  this.bounds.size.height / 2 * fontScale;
+   // if(text.bounds.size.width > this.bounds.size.width){
+   //   text.fontSize =  this.bounds.size.width / 3 * fontScale;
+   // }
     text.position.x = this.width / 2;
     text.position.y = this.height / 6;
     this.text.addChild(text);
@@ -465,10 +465,10 @@ Label.prototype.placeText = function(){
     text.content = leftTextB;
     text.justification = 'center';
     text.fillColor = 'black';
-    text.fontSize =  this.bounds.size.height / 2 * fontScale;
-    if(text.bounds.size.height > this.bounds.size.width / 2){
-      text.fontSize =  this.bounds.size.width / 2 * fontScale;
-    }
+    text.fontSize = 10; //  this.bounds.size.height / 2 * fontScale;
+   //  if(text.bounds.size.height > this.bounds.size.width / 2){
+   //    text.fontSize =  this.bounds.size.width / 2 * fontScale;
+   //  }
[…]

### 2021-11-04 finalproject/7x7/7x7.tex
added images

\documentclass[11pt]{report}
\usepackage{geometry}
\geometry{letterpaper, top=1in, bottom=1in, left=1in, right=1in}                                 
\usepackage[parfill]{parskip}   
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{array}
\usepackage{tabularx}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan
    }

\setlength{\parindent}{0em}
\pagenumbering{gobble}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\newday}[1]{\newpage{\LARGE \textbf{#1}}\\}
\newcommand{\imagefolder}{/Users/lauriaclarke/Documents/mfadt/ms1/lauriaclarke.github.io/instructionsetsforstrangers}

\title{{\LARGE\textbf{7 in 7}} \\ (all the photos are in the images directory, just haven't had time to place them properly in the document)}
\date{\today}
%\author{}

\begin{document}
\maketitle
%\tableofcontents
%----------------------------------------------
\newpage
\subsection*{five (11/03/2021)}
\textbf{name:} trachea support trial no. 1

\textbf{soundtrack:} Ice On The Dune by Empire of the Sun

The objective of this prototype was to create a mechanically robust way to actually make the trachea prototype move and such. First, I did some sketching and head scratching. Then I felt overwhelmed, but I kept thinking. I decided I needed a support on both the curved and flat sides of the trachea tube. One side would just act as support, while the other side allowed the monofillament to be pulled, causing the trachea to contract. After thinking through all kinds of unweidly solutions I had the idea to use pieces of dowel with small holes drilled through both to anchor the trachea and provide the pulling surface.

So I put my shoes on and walked the 80 feet between my apartment and the hardware store -- this apartment is perfect in so many ways -- and got a 3/8" dowel for \$1.40. I measured out a length long enough to suspend the trachea and sit inside a piece of scrap wood and cut two pieces.

Then I marked out where the holes should go; just in line with the vinyl tubing in the trachea. Next, I carefully punched and drilled some small holes through the dowel and added some 3/8" holes -- with a few different spacings -- to my scrap wood to seat the dowels.

The next step was tricky. I needed to create a way to anchor the curved side of the trachea to one of the dowels with holes in it. The catch being no glue was allowed (I alwyas try hard not to use glue when prototyping for maximum material resuse). After thinking for a while, I took a sewing needle and threaded it with monofilament. Then, I carefully pierced and went through the back of the curved side of each vinyl tube leaving about 4" of monofillaent on both sides of the hole.

Finally, I was ready. Starting with the curved, anchor side, I attached one of the dowels to the trachea. Then I seated the anchor dowel in my scrap wood and threaded the other side through the holes. Tying knots in monofillament is always fiddly. The final result looked pretty cool

Then I got to play with it! Wow, it worked better than anticipated!

\begin{figure}[h]
\centering
  \includegraphics[width=0.22\textwidth]{images/46.JPG}
  \includegraphics[width=0.22\textwidth]{images/47.JPG}
  \includegraphics[width=0.22\textwidth]{images/48.JPG}
  \includegraphics[width=0.22\textwidth]{images/49.JPG}
  \includegraphics[width=0.22\textwidth]{images/50.JPG}
  \includegraphics[width=0.22\textwidth]{images/51.JPG}
  \includegraphics[width=0.22\textwidth]{images/52.JPG}
  \caption{sequence of contracting each ring then releasing all of them slowly}
\end{figure}

Takeaways:
\begin{itemize}
\item
it's important to use a sharp needle when going through latex cause it's really stretchy
\item
I need to get this in a position so I can move air through it!
\item
  this works really well for something that's 3" long, but I'll need to use something that's much more ridgid for the 
\end{itemize}
\clearpage
\subsection*{four (11/03/2021)}
\textbf{name:} latex / trachea trial no. 2

\textbf{soundtrack:} Gillian Welch discography all day (every couple months I spend one day listening to all her music)

The intent of this prototype was to see how close to a trachea structure I could get. I repeated the same experiemnt from Tuesday night more carefully and with more components in a nattempt to construct ad better trachea protptype. The goal was to create a closed tube with a mechanism in place to create a change in shape.

First I carefully cur my pieces of latex to size -- two 3" x 1.5" pices. Then I cut four pieces of vinyl tubing to 1.5" lengths to embed in the latex. I cleaned and prepared the latex with glue, then sandwiched the vinyl tubing between the latex pieces. I used a pretty thin and even layer of glue and was very careful to prevent air bubbles.

Next I took another 1.5" wide piece of latex and applied glue to the edge. I applied glue to one edge of my tubing assembly and after waiting the appropriate aount of time, carefully stuck the two edges together with about 1/4" of overlap.

The next step -- closing the tube -- was tricky. To prevent the vinyl tubing from pulling apart the new seams I took pieces of wire and threaded them through the tubing. Then I bent them into a U shape keeping the twp sides I waas gluing close together with little tension on them. I then applied glue to the other side of the sandwich andt he other isde of the flap and sealed the other side.

After letting this sit for some time, I began to remove the wire pieces and replace them with bits of monofilament like in the previous prototype. I was pleasantly surprised by how well the glue sustained the tension from the vinyl tubing. Nothing began to tear or separate sinificantly, though there was one spot that looked like it didn't get enough glue.

Things I learned:
\begin{itemize}
\item
  this was pretty a easy easy with a 3" section, but I think it will be more difficult with a longer tube
\item
  the latex to latex bond is pretty strong and didn't come apart from the pressure of the viny tubing
\item
  I need to think of a clever way to mount this so that it can be mechanised
\item
  I ended up skimming a cool medical paper about breath noises while making this prototype...need to revisit BUT the gist was that the brachea and trachea are the really important parts of breath noises. The coolest thing I learned -- aside from a bunch of laminar flow stuff in the lungs -- is that our organs and chest stuff act as a low pass filter for noises coming from our lungs and lower brachea / trachea. Most of the higher frequency breath noises that we hear come from the neck where there's less tissue for the sound to travel through. Waves! Definitley need to read more thoracic medicine papers.
\end{itemize}

\clearpage
\section*{nothing (11/02/2021)}

Had to do work work for my job instead.

\clearpage
\section*{three (11/01/2021)}
\textbf{name:} latex trial no. 1

\textbf{soundtrack:} my own breath in a respirator

This prototype is also going to span multiple days and build on itself. The intent of this first try was to get a feel for working with latex.

First I assembled my materials. A few 2" x 2" squares of 14 mil latex, some rubbing alchohol, and a bottle of rubber cement. Then I prepared my work area, which was a little tricky since the window in my studio wouldn't open. I ended up working on top of the stove with the vent hood on while weraing a respirator.

After reading a little bit on the internet I went for it. The order of operations is to put glue on both pieces of latex, let it dry for about 5 minutes and then stick the two together. When you put the glue on initially the latex shrivels up. As time passes it relaxes and returns to its original shape. Even though I expected this I was pretty surprised to see just how much it shrivled. I wasn't really sire about the rubbing alchohol so I left it out at first.

The first test I did turned out well, all things considered. The two pieces of latex were firmly stuck together. While the latex was relaxing I started watching a youtube video for more instruction. Even though I hate instructional youtube videos, this one was quite helpful. I learned that I was using too much glue and that it's a good idea to wipe the latex with rubbing alchohol before applying glue, as I suspected.

For the second test I used a q-tip and first wiped the gluing area with isopropyl. Then, using another q-tip I applied a much thinner layer of glue than before. The latex still curled up, but it was less messy and uncurled faster. I was also more mindful to avoid air bubbles as I stuck the two pieces together. This test went pretty well, but still had a lot of air bubbles and wasn't very flat.

For the last test I decided to embed two small pieces of vinyl tubing between the sheets of latex to play with changing the shape of the sheet. I applied the glue as usual and then once it had dried set the two pieces of tube on the glue side oasdf one piece. I then carefully applied the other piece of glue latex. This time I was veyr mindful not to get any airbubbles and the final result was quite smooth. Its appearnce is certainly reminiscent of ridged tissue with cartalidge.

Then I threaded some monofilament through the tube and some glass beads I found in the dryer with my laundry -- the fact that I can always seem to find the materials I need, but not my fancy new bike sock is irksome but feels fair -- and played around. Without a structure to hold the tube it was tricky to contract it.

Things I learned and new questions to address:
\begin{itemize}
\item
the british guy on youtube said you should only work with a small vessel of gue and isopropyl at a time to minimize evaporation...need to find some tiny jars
\item
need to determine how strong the bond is -- how could I create an anchor point needed to mechanize the diaphragm?
\item
my experiement with the vinyl tubing is cool, but it would be better if the vinyl formed a closed tube and the whole thing was more mechnically robust
\item
what sounds or feeling does water have on the latex?
\item
I continue to have a slight allergy to latex powder. Hopefully it doesn't get worse. I will find nitrile gloves and continue not to touch my face.
\item
...more questions to come
\end{itemize}

\clearpage
\section*{two (10/31/2021)}

\textbf{name:} disembodied breath no. 1

\textbf{soundtrack:} some of Mirrors by Angel Olsen and Late Night Feelings by Mark Ronson et. al ...piecemeal listening today

This prototype is defintely going to span two days. Today being Halloween seemed like an appropriate day to start, however.

The goal of this prototype is to understand the response to the sound of disembodied breath. To achieve this will use my trusty buckets to "produce" breath sounds. I would say that this prototype is Wizard of Oz adjacent.

Fist I took a small bluetooth speaker and the biggest USB battery I have. I made sure that they fit inside one 5 gallong bucket when another was placed on inside.

Then I put the bucket from yesterday with the hole and latex tube attached inside the bucket with the speaker.

I connected to the speaker with my cellphone and played a youtube video of breath noises at a loud volume. It was a very striking effect.

I took my buckets -- the new, hottest trend in streetwear accessories -- out on the street and found a cafe on a pretty busy corner. I put them on the sidewalk and played breath noises.

A number of things went wrong; I needed to use my cellphone for something other than youtube, the bluetooth kept disconnecting, and due to street noise and music from the cafe no one really heard the bucket. I also placed it near a trash can and was worried someone would throw something yucky inside.

Then I went home and created an mp3 from the youtube video. I put it on an old cellphone on repeat and used an aux cable to connect the speaker to the cellphone.

Sometime later I took this, much more solid, combination out to my favorite corner bar with my sister who was visiting. And placed it near a crosswalk (the same one I used for instruction sets for strangers, in fact).

Not too many people noticed the bucket. I think the road noise was a lot to contend with. The people that did notice were pretty funny. Three interactions stand out:
\begin{itemize}
\item
A young woman kept looking at the bucket while waiting for the crosswalk with a group of friends. She kept looking furtively over her shoulder at the bucket and was clearly perturbed by the noise. It was also clear from her interaction walking away that she didn't immediately mention it to her companions. 
\item
A man in his late 50s perhaps was eating a slice of pizza while crossing the street. When he got to the bucket he stopped and was very curious. He looked inside and listened closely. He clearly wasn't in a rush as he finished his pizza. Throughout the interaction his facial expression was a mix of confusion, disgust and laughter. He walked away shaking his head. 
\item
A young man was smoking a cigarette near the bucket. After some time he started looking at it more closely and clearly noticed the sound. He seemed confused and a bit disconcerted. My sister and I were clearly looking at him so he came over to us and mentioned that he though someone had ``planted something" on the corner. He said he'd thought it was music and had been waiting for a pop song to start playing, but nothing happened. 
\item
Finally, a woman collecting recycling walked by. She stopped at the trash can adjacent to the bucket. And then I watched in seeminly slow motion as her attention turned to my prized buckets. She seemed a bit confused, but bent over and took the inner bucket out. At that point I left my seat and went over and asked her to stop. I explained that the buckets were mine and she apologized and went on her way. Then it was time ot go home.
\end{itemize}

The big takeaways which I want to adjust on the next iteration aere as follows:
\begin{itemize}
\item
needs to be louder
\item
needs a more quiet location
\item
need a better way to control which breath sounds are playing
\end{itemize}

To account for these, I think I'm going to try a park. I can't find a louder speaker, but maybe I can make the noise louder digitally. I am also going to create simple interface which will let me adjust which breath sounds are playing instead of having to listen to the youtube video mp3 on repeat.
\clearpage
\subsection*{one (10/30/2021)}

\textbf{name:} diaphgarm test no. 1

\textbf{soundtrack:} Treehouse by Sofi Tukker

The objective of this prototype was to create a mechanism for forcing air through a tube so that I can play with different sounds. To do this I intended to create a diaphragm to move the air. My intention was to attach a pice of latex over one end of the bucket and a small tub at the other.

I started with one of the 5 gallon buckets I found and bought a bucket lid (should've tried harder to find the lids when I found the buckets).

I cut the center out of the lid and then did a test to see how well the latex stayed in place under the lid. Luckily it worked well and didn't damage the latex.

Next, I drilled a hole in the bottom of the bucket. I forgot about my hole saw kit and just used a knife to slowly shave off the edges until it was the right size for the tube.

Then I trimmed a pice of Latex to size (this required a fresh box cutter since all my other knives were too dull for a clean cut).

Finally, I attached the latex to the open end of the bucket under the "lid" and poked the tub through the hole in the bottom.

The effect was pretty immediate. By depressing the diaphragm I easliy forced air out through the tube -- an exhale -- and air came back into the tube when I let the diaphragm rebound -- inhale.

There were three pnysical properties I wanted to play with to see how they impacted the sound coming from the prototype:
\begin{enumerate}
\item
how much of the tube was outside of the bucket vs. inside
\begin{itemize}
\item
This didn't seem to have a big impact. It was really ahrd to tell the difference between the tube being all the way out and all the way in. I think I need to cut the tube ot be different lengths.
\end{itemize}
\item
how sound changed when the tube was pinched in different places
\begin{itemize}
  \item
    when the tube was pinched near the bottom the sound was lower in pitch
  \item
    when the tube was pinched near the top the sounds was higher in pitch
  \item
    nearer the top sounded more like human breathing
\end{itemize}
\item
how sounds changed if the bucket was crushed sideways to be less round
\begin{itemize}
  \item
    this one was also hard to tell, it seemed that there was slighly less resonance when the bucket was crushed
  \item
    the intent was to make the sound less hollow and I think this was slightly true
\end{itemize}

\end{enumerate}

\vspace{1cm}
I think this prototype was pretty successful. It left me with a good number of questions which are as follows.
\begin{itemize}
\item
  how does the length of the output tube effect the sound? need to experiement, but an not ready to cut the tube.
\item
  how does the shape of the cavity effect the sound?
\item
  is this cavity ridiculously large in comparison to normal lungs?
\item
  how do I make the sound less hollow?
\item
  what volume of air is displaced? (if I had a flow meter I could just integrate, but I'll have to be more clever)
\item
  how would a lung structure change the sound?
\item
  how would the sound change if there was some kind of foam or cloth inside the cavity? would that sound less hollow?
\item
  how the heck am I going to automate the diaphragm?
\end{itemize}

\clearpage
\section*{7 in 7 planning}

The first step of any prototyping exercise is to collect materials. I did some of this a couple weeks ago when I went to canal rubber (so cool, so much sneezy latex powder). This was good fodder for the diaphragm and trachea prototypes, but I still need a chest cavity like thing that would be more robust than the small plastic container I tried already. A large plastic carboy for brewing would be ideal, but that violates my prototyping rule of only buying things you truly need for a material or functional evaluation (like latex) and can't find anywhere else. While out on a late walk this Monday evening I found a lot of recycling outside a restaurant near my apartment. Isolated inside a single clear trashbag were three 5 gallon buckets formerly containing pickles. These streets deliver! They didn't even smell much like pickles once I got them home and opened the bag. After a few days outside and a big rainstorm they were good to go. Seems like I already have the necessary materials for my other physical protptypes.

\subsection*{proposed prototypes}

There are two sensory aspects of this work, the auditory and the physical or visual. It was pretty easy to identify prototypes that fell in each of these categories. The auditory prototypes should play with differnet types of breath and evaluate the effect of disembodied breath. The physical prototype should focus solely on the mechanical aspec of reproducing breath according to the basic breakdown of the respiratory system I've outlined. I'm not going to start thinking about the electronics used to actuate the thigns.
%outside of investigating what might be quiet.

It was less easy to think about prototypes that address the less tangible questions I am hoping to probe with this work. To think about this I started by looking back at the questions I asked myself originally. After reviewing that list, the big question I want to start asnwering is how a viewer sees their own presesnce near the work impacting its breath. This is an answer to the questions of who holds power over our own breath and how far does breath extend as a metaphor for out relationship with the natural world. I'm going to start with this one isolated question because it has a lot of facets and ultimately will lead to the bigger question of what should the viewer take away from the work (in other words, what experience do I intend for them to have).

Based on the above thinking I concoted the following protptypes.

%What if as more people enter a space, or come closer to the contraption, it becomes more difficult for it to breath?

\vspace{1 cm}
\textbf{interactive (how a viewer sees their own presence near the work impacting its breath):}

as someone approaches the bucket the breath becomes worse

as someone leaves the bucket breath becomes worse

breath eminates from a tree? that's a bit random, but perhaps illustrative

\vspace{1 cm}
\textbf{auditory (addressing disembodied breath, how breath denotes sickness, difficulty, and emotion):}

Hide a speaker and battery inside two 5 gallon buckets in a public place with moderate traffic and see how people react. Do each of these different types of breath illicit different reactions?

play normal breath sounds

play laboured and sick (covid equivalent) breath sounds

play emotional breath sounds (sad, happy, aroused, etc.)

\vspace{1 cm}
\textbf{visual / physical (addressing how to actually make this thing and how it should look):}

make a big diaphragm and play with \textit{quiet} ways to make it move

make a trachea protptype with multiple rings that contract

play with the idea of a larynx

model a brachea in openSCAD and print one small piece to think about how moitsure could interact with this piece

\end{document}

### 2021-11-05 finalproject/report/finalproject.tex
added report to github

--- 2021-11-04 finalproject/7x7/7x7.tex
@@
 
-\title{{\LARGE\textbf{7 in 7}} \\ (all the photos are in the images directory, just haven't had time to place them properly in the document)}
+\title{Major Studio I \\ Final Project}
 \date{\today}
-%\author{}
+\author{Lauria Clarke}
 
@@
 %----------------------------------------------
+\section*{the idea}
+
+The Nature of Breath is a machine that breathes where the speed, depth, and "wetness" of the breathing varies in response to COVID 19 data.
+
+Since COVID started we have all experienced a global fascination with breath.
+
+It should be naturalistic in the auditory component and artificial in the visual component.
+
+\section*{mechanical breakdown}
+
+Thinking about the way this device will function mechanically is the hardest part since there are few --- if any -- existing examples.
+To help think about the systems in a more compartmentalized way, I have broken it wodn into three sub systems: actuation, storage, and modulation.
+Actuation is the mechanism taht forces air into or outof the system, storage is where air is stored between breaths, and modulation is the subsystem which changes the flow of air and produces the appropriate sound.
+
+In the human body, these three pieces combine to produce breath.
+
+\subsection*{actuation}
+
+In the body, the thing that actuates airflow through the respiratory system is the diaphragm. It contracts causing air ot be firces in or out of the lungs.
+
+Actuation could be achieved in a two ways in this project.
+First, a changing pressure differential forces air in and out of storage medium.
+This could be accomplished by changing the pressure in a sealed volume which contains the storage comoponent by either using a diaphragm or directly changing the pressure using a pump.
+Second, the actuation could be direct -- a bi-directional pump forces air through the airway without any kind of storage component.
+This solution reduces complexity, but may increase cost.
+
+Either way, the acuation technique needs to be bidirectional -- inhale and exhale.
+
+\subsection*{storage}
+
+Storage isn't necessary if actuation is direct on the airway. For visual purposes however, it would be quite striking to see a shape inflating and deflating with each breath.
+
+\subsection*{modulation}
+
+Modulation is the most complicated and exciting part of this project.
+It can be furthe briken down into the trachea, larynx, and mouth/sinuses.
+Each of these three parts of the respiratory system have an impact on how our breating sounds.
+
+In this system, the trachea will be the first point of focus.
+The trachea consists of C shaped pieces of cartiladge stack on top of each other with a smooth muscle closing the C.
+When the smooth muscle contracts the trachea grows more narrow allowing us to do things like cough.
+
+% include image of trachea here
+
+This mechanism could pretty simple to replicate.
+
+\section*{initial material gatering and prototypes}
+
+I was quite consumed by this idea after having it, but my ideas for implementation were too flimsy and became quite consuming.
+I called a number of places that sell balloons and 5 gallon water jug hoping ot find tha materias needed to replicate a gigantic version of that gradeschool experiment.
+Didn't have much luck.
+I still couldn't stop thinking about all this, however, so I went to Canal Rubber Co.
+to look at sheets of latex.
+This was a GREAT idea.
+It was soo cool!
+I gathered more materials than I actually needed and probably spent more than I should've -- \$30.
+
+Then I got home and started playing with the materials and rememebred I'm a bit allergic to latex powder. Ooops!
+
+I build a simple diaphragm test.
+
+\clearpage
+\section*{midterm response}
+
+Hi Harpreet, I neglected to read the prompt very thoroughly for this assignment, so what follows is a very unpolished assesment of my current thinking here.
+More of a journal entry than anything else.
+My apologies.
+
+After being here (in MFADT) for a couple months I belive I am persuing the following learning goal: to know when to engage in blindly intuitive making and when to engage in research based, highly considered making.
+So far my life has revolved around intuitive making, an inexplicable itch.
+Making first and assigning meaning second.
+I am here, however, to be challenged and that challenge is to become intentional.
+
+On Wednesday, Franco and I went out and had a beer after class.
+It felt a bit irresponsible from a time management perspective, but it was really helpful to talk about the midterm process totally outside of school while it was so fresh.
+The bulk of our conversation focused on language.
+One aspect of Franco's presentation which I really appreciated was the framework he developed to decribe what he was discussing.
+For me, language is a big focus right now as I think about what it means to be intentional in my work.
+In order to be intentional we need to posses the language to describe our intent.
+Finding the appropriate language to use to describe what I'm seeing and trying to achieve feels like the next, most important step in my process.
+
+The easiest place to look for language is in what others have already written.
+I spent some time looking for this during the year before I came to school, but without knowing the languae to look for and use in my search it was difficult to find a path of work to follow.
+This was frustrating, because I'm usually pretty good at self directed research and reading.
+Anyway, I'm definitely in the right place now with the right people to guide me in asking questions and building a foundation of thought for finding answers.
+Language allows you not only to describe, but question a field.
+
+On the topic of building foundation, I spoke with Tega Brain two weeks ago.
+I was extremely nervous and practiced a lot.
+It was a great conversation and she was very generous with her time and knowledge.
+(As and aside, I am so incredibly excited and grateful be in this place where everyone is willing ot speak with me and discuss the things I'm interested in!
+I have felt out of place for so much of my academic and professional life it is wonderful to leave taht feeling behind.) I first asked her what frameworks she had encoutnered over her career which have helped lend more specific language and structure to the way the natural and manmade worlds are described and designated.
+She provided me with many useful references that address that broad ideas I'm trying to learn about.
+
+The conversation naturally flowed into the next question I wanted to ask which was about the similarities between natural optimization processes and manmade or technological optimization procesed.
+We agreed that the scale of time and the objective of optimization tend to be the distinguishing factors between the two.
+In the same vein, she described an overarching urge in her work to replicate manmade systems in natural ways, often relying on this idea of optimization.
+It was helpful to hear that specific sense of purpose stated in such plain language.
+
+Finally I asked her about the feeling of uncertainty which I am trying to investigate which I fell exists in a lot of her work.
+She asked if I was trying to describe the uncanny valley.
+I don't think I am, at least not ``the'' uncanny valley.
+I believe that the unvanny vally is an appropriate technique to describe what I'm talking about, it is accurate in reflecting the feeling of unease, but is specifically describes a human centric condition.
+What about more broad, natural existence?
+That is what I'm after.
+Maybe someone with a better understanding of this question will tell me that the question itself is unnecessary and has already been answered, but that's for me to find out on my own.
+
+Thinking a bit about the review itself, I found the process of preparing the presentation much more helpful than delivering the presentaion.
+Feedback was certainly useful, but I think I could've structured my presentation to make it more helpful.
+By framing a few specific questions for the audience to ponder I could've directed the feedback.
+Definitely need to rememebr that next time.
+Coming up with questions for others to asses for you impllies that know what help you need with so that you can ask for it.
+A little circular by definition.
+
+I really liked mark's suggestion of framing the project in diferent ways to understand it better.
+I'm not sure I entirely know how to go about doing that, but I'm definitly going to try.
+Richard's coments about following my intuition toward this vein of inquire was helpful in that it made me think more specifically about the sometimes conflicting roles of intuition and intention in this process.
+The thoughful feedback from class was overwhelming in how geenrous it was (again, feeling lucky to be here)!
+In many of my classmates' comments I head the same question, which Tanve said really well: what do you want people to take away from your work?
+Again, a suggestion to focus on my intent.
+
+Overall, the focus on building research and background before engaging in making has made me a lttle bit wary of my own ideas.
+I think this new degree of consideration is good, but I also recognize that it's important to let an urge to create drive your forward momentum.
+As I grow as thinkiner and come to understand the space I'm trying to exist in, my process will naturally become more considered.
+But I've got to start somewhere, and that's right now with the excitement and ideas I have in this moment.
+
+There is a lot of reading ahead of me. Below is an accumulated list from the last month or so.
+
+\section*{accumulated reading list and internet suggestions}
+
+Biomimicry - Janine Benyus (my dad...who has a weird knack for giving me important books at critical moments, like the Toaster Project right when it came out during my senior year of highschool)
+
+seeing is forgetting the name of the thing one sees - Lawrence Welschler (Ed Andrews)
+
+Anthropocene Thesis (not actually the name) - Donna Haraway (Tega Brain)
+
+Against the Anthropocene - T. J. Demos (Tega Brain)
+
+Exhallation - Ted Chang (Tega Brain)
+
+The Sciences of the Artificial - Herbert Simon (Mark ?)
+
+Life's Edge - Carl Zimmer (me)
+
+Making and Being -
+
+\vspace{1cm}
+Internet followups from Tega Brain:
+
+\url{http://www.greyroom.org/issues/68/72/the-smartness-mandate-notes-toward-a-critique/}
+
+\url{https://alanwarburton.co.uk/fairytales}
+
+\url{https://jods.mitpress.mit.edu/pub/lewis-arista-pechawis-kite/release/1}
+
+\url{http://www.madlab.cc/}
+
+think this one is the donna haraway\\
+\url{https://www.e-flux.com/journal/75/67125/tentacular-thinking-anthropocene-capitalocene-chthulucene/}
+
+\vspace{1cm}
+Internet followups from Harpreet:
+
+look at ISEA 2021
+
+\url{https://www.ianingram.org/machines/2009_squirrel.html}
+
+\url{https://www.google.com/search?q=bodies+in+motion&rlz=1C5CHFA_enUS719US719&sxsrf=AOaemvI9jBtAv206EcJ4FUexQurm74VNBw:1633968526778&source=lnms&tbm=isch&sa=X&ved=2ahUKEwiJ7vKM38LzAhXrQ98KHbZjBGoQ_AUoAnoECAIQBA&biw=1440&bih=766&dpr=2}
+
+\url{https://vimeo.com/220895476}
+
+\url{https://news.mit.edu/2018/creating-3-d-printed-motion-sculptures-from-2-d-videos-mit-csail-0919}
+
+\url{https://vvvv.org/blog/bodies-in-motion-humanscale-milan-design-week-2019#:~:text=Bodies%20in%20Motion%20is%20an,scientific%20approach%20to%20furniture%20design}
+
+\url{https://www.youtube.com/watch?v=_v3J6D6R_nI&ab_channel=HumanscaleHQ}
+
+philip beesley
+
+\url{https://www.christianhubert.com/writing/assemblage}
+
+\url{https://link.springer.com/chapter/10.1057/9781137025012_4}
+
+\vspace{1cm}
+Internet followups from Richard The:
+
+Design Triennale, Broken Nature
+
+RadioLab -- Breathing podcast
+
+\url{https://www.lozano-hemmer.com/last\_breath.php}
+
+\vspace{1cm}
+from Guin: \url{https://www.asianstudies.org/publications/eaa/archives/history-and-sustainability-of-bunraku-the-japanese-puppet-theater/}
+
+from Tanve: The Year the Earth Changed - David Attenborough
+%random 
+%Building a foundation of language upon which to base your work is critical.
+%
+%Before coming ot Parsons my practice was veyr much in the vein of making first and assigning meaning second.
+At present I feel as though I'm being pushed to be more intentional in all areas of my making.
+%
+%
+%BUIlding a framework or background of language also allows you to call upon references or the prior work in your field more easily. 
+%
+%I am missing the critical theory that supports the more intuitive exploration I've done so far.
+I have been looking for it, but without knowing the languae to use in my search it has been difficult to find on my own.
+A good reminder too that language allows you not only to describe, but question a field.
+%
+%I value the feedbak
+%
+%talk about approaching the work from multiple angles
+%I have a hard time letting go of my desire to solve physical challenges through making.
+This particular desire has very deeply rooted in me for as long as I can remember.
+Coming up with a physical solution to a complex and often unecessary problem is what drive a lot of my work.
+%
+%Again, I think that I need to asses this urge through the lens of intentionality.
+It can be useful at time, but learning somplementary ways of thinking about the intent or goal of my work is a skil which I really hope to develop in my time here at parsons.
+%
+%I spoke with tega brain last week and that was a fantastic conversation.
+She provided me with many, many references that address the questions in her own work and mine.
+The conversation we had about the specific questions I asked her felt really fruitful as well -- I should've typed this up sooner.
+We spoke about the paralles between optimization in nature and optimization in engineering and more.
+%
+%
+%chat with Tega Brain
+%
+%chat with Rob
+%
+%
+%When Mark asked my why ths question of defining what is natural vs.
+artificial matters I felt like I stumbled a bit.
+The answers that usually run through my head felt wrong for the moment, but I need to get better at saying what I mean.
+I have this idealized scenario where a kid grows up in a house where the deck outside is made of fake wood and all the floors in the house are made of fake wood and every single piece of wood around is actually plastic - exactly like Through the Arc of the Rainforest(!).
+This kid never know what real wood is.
+Wouldn't that be problematic?
+To be so out of touch.
+Maybe I need to re-read through the arc of the rainforest for better insight on this one.
+Although everyone dies at the end of that book so maybe it's not really so helpful.
+%
+%
+%
+%
+%suggestion from tanve: what do you want people to take away from your work?
+
+\clearpage
+\section*{chat with harpreet on 10/25}
+
+7 in 7 is an attempt to try multiple things in many ways
+
+use different perspective to address the same question
+
+could use
+
+is the conecptual prototype having the intended effect on users
+
+are the users having the intended experience
+
+\clearpage
+\section*{{\LARGE7 in 7} (in reverse order, accompanying photos forthcoming)}
+also, spellcheck in my text editor isn't working for this document
+
 \newpage
@@
 
-\textbf{soundtrack:} Ice On The Dune by Empire of the Sun
+\textbf{soundtrack:} Empire of the Sun
 
-The objective of this prototype was to create a mechanically robust way to actually make the trachea prototype move and such.
-First, I did some sketching and head scratching.
-Then I felt overwhelmed, but I kept thinking.
-I decided I needed a support on both the curved and flat sides of the trachea tube.
-One side would just act as support, while the other side allowed the monofillament to be pulled, causing the trachea to contract.
-After thinking through all kinds of unweidly solutions I had the idea to use pieces of dowel with small holes drilled through both to anchor the trachea and provide the pulling surface.
-
-So I put my shoes on and walked the 80 feet between my apartment and the hardware store -- this apartment is perfect in so many ways -- and got a 3/8" dowel for \$1.40.
-I measured out a length long enough to suspend the trachea and sit inside a piece of scrap wood and cut two pieces.
-
-Then I marked out where the holes should go; just in line with the vinyl tubing in the trachea.
-Next, I carefully punched and drilled some small holes through the dowel and added some 3/8" holes -- with a few different spacings -- to my scrap wood to seat the dowels.
-
-The next step was tricky.
-I needed to create a way to anchor the curved side of the trachea to one of the dowels with holes in it.
-The catch being no glue was allowed (I alwyas try hard not to use glue when prototyping for maximum material resuse).
-After thinking for a while, I took a sewing needle and threaded it with monofilament.
-Then, I carefully pierced and went through the back of the curved side of each vinyl tube leaving about 4" of monofillaent on both sides of the hole.
-
-Finally, I was ready.
-Starting with the curved, anchor side, I attached one of the dowels to the trachea.
-Then I seated the anchor dowel in my scrap wood and threaded the other side through the holes.
-Tying knots in monofillament is always fiddly.
-The final result looked pretty cool
-
-Then I got to play with it! Wow, it worked better than anticipated!
-
-\begin{figure}[h]
-\centering
-  \includegraphics[width=0.22\textwidth]{images/46.JPG}
-  \includegraphics[width=0.22\textwidth]{images/47.JPG}
-  \includegraphics[width=0.22\textwidth]{images/48.JPG}
-  \includegraphics[width=0.22\textwidth]{images/49.JPG}
-  \includegraphics[width=0.22\textwidth]{images/50.JPG}
-  \includegraphics[width=0.22\textwidth]{images/51.JPG}
-  \includegraphics[width=0.22\textwidth]{images/52.JPG}
-  \caption{sequence of contracting each ring then releasing all of them slowly}
-\end{figure}
-
-Takeaways:
-\begin{itemize}
-\item
-it's important to use a sharp needle when going through latex cause it's really stretchy
-\item
-I need to get this in a position so I can move air through it!
-\item
-  this works really well for something that's 3" long, but I'll need to use something that's much more ridgid for the 
-\end{itemize}
-\clearpage
+\newpage
 \subsection*{four (11/03/2021)}
@@
 I was pleasantly surprised by how well the glue sustained the tension from the vinyl tubing.
-Nothing began to tear or separate sinificantly, though there was one spot that looked like it didn't get enough glue.
+Nothing began to tear or separate.
 
 Things I learned:
-\begin{itemize}
+\begin{itmemize}
 \item
@@
 \item
-  I ended up skimming a cool medical paper about breath noises while making this prototype...need to revisit BUT the gist was that the brachea and trachea are the really important parts of breath noises.
-The coolest thing I learned -- aside from a bunch of laminar flow stuff in the lungs -- is that our organs and chest stuff act as a low pass filter for noises coming from our lungs and lower brachea / trachea.
-Most of the higher frequency breath noises that we hear come from the neck where there's less tissue for the sound to travel through.
-Waves!
-Definitley need to read more thoracic medicine papers.
+
 \end{itemize}
 
-\clearpage
-\section*{nothing (11/02/2021)}
+\newpage
+\subsection*{nothing (11/02/2021)}
 
@@
 
-\clearpage
-\section*{three (11/01/2021)}
+\newpage
+\subsection*{three (11/01/2021)}
 \textbf{name:} latex trial no. 1
@@
 
-\clearpage
-\section*{two (10/31/2021)}
+\newpage
+\subsection*{two (10/31/2021)}
 
@@
 I am also going to create simple interface which will let me adjust which breath sounds are playing instead of having to listen to the youtube video mp3 on repeat.
-\clearpage
+\newpage
 \subsection*{one (10/30/2021)}
@@
 \end{itemize}
-
-\clearpage
+\newpage
 \section*{7 in 7 planning}

### 2021-11-12 ccgame/p7_sketch.js
fixed a bug in my game

@@ -152,6 +152,7 @@ function resetScenes(){
   for(let i = 0; i < scenes.length; i++){
     scenes[i].isDone = false;
     scenes[i].introDone = false;
+    scenes[i].resetThings();
   }
 }
 

### 2021-11-12 ccgame/scene.js

@@ -123,6 +123,12 @@ Scene.prototype.build = function(){
   }
 }
 
+
+Scene.prototype.resetThings = function(){
+  for(let i = 0; i < this.nThings; i++){
+    this.things[i].reset();
+  }
+}
 //----------------------------------
 // CUSTOM DRAW FUNCTIONS
 // intial scene of the game

### 2021-11-12 ccgame/thing.js

@@ -55,6 +55,14 @@ Thing.prototype.update = function(){
   }
 }
 
+// update the activation of the thing
+Thing.prototype.reset = function(){
+  // if the activation count == number, thing is activated 
+  this.actCnt = 0;
+  this.activated = false;
+  this.completed = false;
+}
+
 // print a messgae about the item when we get close and interact with it
 Thing.prototype.printMessage = function(){
   push();

### 2021-11-15 misc/biodesign.tex
misc updates as well as prototype one presentation

\documentclass[11pt]{report}
\usepackage{geometry}
\geometry{letterpaper, top=1in, bottom=1in, left=1in, right=1in}                                 
\usepackage[parfill]{parskip}   
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{array}
\usepackage{tabularx}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan
    }

\setlength{\parindent}{0em}
\pagenumbering{gobble}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\newday}[1]{\newpage{\LARGE \textbf{#1}}\\}
\newcommand{\imagefolder}{/Users/lauriaclarke/Documents/mfadt/ms1/lauriaclarke.github.io/instructionsetsforstrangers}

\title{Major Studio I \\ Final Project}
\date{\today}
\author{Lauria Clarke}

\begin{document}
\maketitle
%\tableofcontents
%----------------------------------------------
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              \tx
\textbf{Why you are interested in taking this class and participating in the Biodesign Challenge?}                                                                                                                                                                                                                                                                                                                                                                                                                                                   
I would like to take this course and participate in the biodesign challenge because I am trying to build a foundation of thought in the areas of biodesign and bioart. My prior background is in computer engineering and kinetic sculpture and I have always been fascinated by the field of biomimetics.

I am perpetually drawn to this field because of the ambiguity and contrast it highlights -- blurring the line between what we intuitively know to be natural and what we intuitively know to be man-made or artificial. Recently, I have been trying to pin down the importance and utility of this particular uncertainty in a broader research contex. The root of my curiosity in addressing the boundary between what is natural and artificial is a desire to understand how creative practitioners can use it to generate reflection and consideration of humanity's place in the natural world.

To conceptualize the impact of human activity on this planet is practically impossible, yet we find ourselves at a juncture where this definition is at the heart of nearly every social and political issue. If Art is a trascendent mechanism for collapsing the scales of Time and space we need it now more than ever to make sense of this recursively enormous question.

While I don't have a specific project idea to contribute to this course right now, I am extremely curious to observe and be a part of the community of designers working at the very heart of these questions. Broadening my vocabulary and foundation of thought to address the above feels critical to the developement of my practice right now.

\textbf{Describe your Science/Biology and Biodesign/Bioart experience or background.}

I am fascinated by naturalistic movement and much of my sculpture based work seeks to mimic naturalistic forms and motion. Outside of the biology I learned during primary and secondary school education, my biology background is entirely intuitive and learned from life spent watching very, very closely while playing outside.

\textbf{If you don’t already have a project and/or team, what are some of the areas of interest you have for investigation, exploration, research and/or intervention/engagement?}

As mentioned above, I am curious to explore the boundary of what we consider to be artificial and natural. Kinetic sculpture is the vocabulary I am comfortable using to address this in my own practice, but I'm eager to colaborate with others and push myself toward new language and modes of inquiry.

-- how do yuo define the recursively enormous --

\end{document}

### 2021-11-18 misc/biodesign.tex
small tweaks

@@
 %----------------------------------------------
-                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              \tx
+
 \textbf{Why you are interested in taking this class and participating in the Biodesign Challenge?}
+
 I would like to take this course and participate in the biodesign challenge because I am trying to build a foundation of thought in the areas of biodesign and bioart.
@@
 Recently, I have been trying to pin down the importance and utility of this particular uncertainty in a broader research contex.
-The root of my curiosity in addressing the boundary between what is natural and artificial is a desire to understand how creative practitioners can use it to generate reflection and consideration of humanity's place in the natural world.
+The root of my curiosity in addressing the boundary between what is natural and artificial is a desire to understand how creative practitioners can use the uncertainty surrouding thie distinction to generate reflection and consideration of humanity's place in the natural world.
 

### 2021-11-28 finalproject/report/ms1report.tex
fixed bibliography

\documentclass[letterpaper]{article}
\usepackage{ms1report}
\usepackage[pdftex]{graphicx}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage[numbers]{natbib}
\pdfinfo{/Title (The Nature of Breath) /Author (Lauria Clarke)}

\title{The Nature of Breath: A Biomimetic Kinetic Sculpture That Breathes}
\author{Lauria Clarke\\ Parsons School of Design, The New School\\ New York, USA\\ clarl404@newschool.edu\\
\newline
\newline
}
\setcounter{secnumdepth}{0}

\begin{document} 
\maketitle

\begin{abstract}
the nature of breath is...
\end{abstract}

\keywords{Keywords}

%------------------------------
\section{Introduction}

% general introduction

Masahiro Mori's 1970 essay prvided the framework for describing the confusing relationship between artifically created human-like robots and, as Mori calls us, real human beings. \cite{mori} This framework of thought has gained significant traction over the last 50 plus years as developments in technology have made it possiblt to mimic the human body with increasing accuracy. This definition has also been critical in the field of computer science as we consinute ot refine mahcine learning techniques which steadily approach the capacities of the uhman brain. This framework describes the point at which an artificial automaton approaches a point of being eeriely similar to a human being, creating a feeling of discomfort via the uncertainty of whether it is "real".

For the lats fifty years this framework fo thought has been critical in describing technology's relationship with humanity.

Over the last *** years however, the fields of biodesign and bioart have increasingly been asking similar questions not about tchnology and humanity, but technology and nature. In light of the increasing environmental pressures of human imposed climate change, these questions are becoming increasingly urgent. Technology is now being used t oreplicate and mimic natural systems. As Donna Haraway noted in her paper, we must thin kbeyond the human. \cite{haraway} The capacity of our technology has surpassed the instinctually introspective and can now mimic what is natural outside of ourselves. This is a critical development and harkens back to mode indigenous ways of thinking - moving and working with natural systems, rather than in opposition to.

If our relationship with technology is extending beyond the human so to must our language. Mori's framework is distinctly human centric. How then do we apply this thinking to non-human natural world? Is Mori's valley merely a snaller depression within a much broader valley encompassing the whole of the natural world? Do humans deserve a place in that broader cintext? To frame this question more broadly, how do we, as human beings, distinguish between what is natural and what is atrificial? How is this distinction shifting as our techological capacity for the artifical increases at great cost to the natural world.

Using the COVID-19 pandemic as lens and reference, this work attempts to lay a foundation for deeping inquiry into the above. While its focus is on a specifically human foram of respiration, respiration in broader context is omnipresent in living organisms, thus providing a jumping off point for larger scope of inquiry into the relationship between what it means to be alive and what it means to be natural. Something about being alive. \cite{zimmer}

\subsection{temporal relevance}
% the pandemic...the soundscape of the pandemic...
Teh COVID-19 pandemic has caused a sort of crystalization on many of the aforementioned optics, and has

During the pandemic many people tried ot mkae ventilators at home. Segue...

\subsection{mechanical relevance}
% talk about the history of breathing apparatus and note that none mimic the human body
Mechanical ventilation has been around for a long time. Since it  seeks to provide "replacement of the respiratory muscles". \cite{ventilatorhistory} This results in apparati that resembly the respiratory system only in the sens that they cause a regutaled chnage in pressure, but have no need for recreating htep rocess by which breaht sounds are actually created. Simialrly, the cield of biomemtic robotics has little use for such inquiry - respiration is not necessary for an electronic humanoid and it is much easier to produce vocal noises in a digital manner.

Two notable works in the field of kinetic sculpture provide reference, Mawa Denkye's lauching machine and another guy's breath machine thing. These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow. %\cite{deynke} \cite{otherguy}

% need to figure out how to combine the two naturally
\section{prior work}

mawa deynke

breath thing from bitforms

homemade ventilators

\section{implementation}

broken down into two steps; thinking and making

\subsection{system breakdown}

the respiratory system is complex, so I made is simpler

\subsubsection{actuation}

how air is moved through the respiratory system

\subsubsection{storage}

how air is stored in the respiraatory system

\subsubsection{modulation}

how the sound of breath is created

\subsection{making}

So far a few different protptypes ahve been attempted. Focusing on the two main areas of inquiry, there have been auditory prototypes, exploring the effect os disembodied breath noises, as well as machnical prototypes, exploring ways to achieve a physiologically accurate breahting mechanism.

\subsubsection{auditory prototypes}

bucket experiement

p5 thing

\subsubsection{physical prototypes}

trachea

diaphragm

lever adjustment

\subsubsection{final prototypes}

the final result was a thing that nearly works and combines many of the elements from above

\section{further work}

there's always more to do...

the boris vian reference

\section{conclusion}

talk about the big distinction of artificial and natural

why it's important to think about this in the context of humanity's relationship with the natural world

%\begin{figure}[h]
%\includegraphics[width=3.31in]{figure.png}
%\caption{This is an example of figure caption. Note that all figures, and tables are to be referenced in the text. \copyright Respect Copyright.}
%\end{figure}
%
%\begin{figure*}
%\includegraphics[width=\textwidth]{two-column-figure.png}
%\caption{Example of a double-column figure with caption. \copyright Respect Copyright.}
%\end{figure*}

\bibliographystyle{plain}
\bibliography{ms1report}

\section{Bibliography}
%The difference between a reference list and a bibliography is that in your references, you list all the sources you directly referred to in the body of your writing in numerical order, whereas a bibliography includes an alphabetical listing of all those authors and sources that you have consulted while writing your essay. Use the same format as for the references otherwise. Those using Latex will follow the usual cite command format.

\section{Author Biography}
Lauria Clarke is an engineer and artist living in New York. She completed an MS in Computer Engineering at Northeastern University in 2017 and is currently persuing an MFA in Design and Technology at The Parsons School of Design. She specializes in digital hardware design and kinetic sculpture and is always curious to learn more about humanity's relationship with the natural world.

% aspires to be a chocolate fish fisherman for Ben and Jerry's.

\end{document}

### 2021-11-29 finalproject/report/ms1report.tex
paper updates, trying ot make a video galleyr

@@
 
-Masahiro Mori's 1970 essay prvided the framework for describing the confusing relationship between artifically created human-like robots and, as Mori calls us, real human beings.
-\cite{mori} This framework of thought has gained significant traction over the last 50 plus years as developments in technology have made it possiblt to mimic the human body with increasing accuracy.
-This definition has also been critical in the field of computer science as we consinute ot refine mahcine learning techniques which steadily approach the capacities of the uhman brain.
-This framework describes the point at which an artificial automaton approaches a point of being eeriely similar to a human being, creating a feeling of discomfort via the uncertainty of whether it is "real".
+Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive.
+While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.
 
-For the lats fifty years this framework fo thought has been critical in describing technology's relationship with humanity.
+% Respiration is an indicator of life at all scales - from single celled organisms to human beings.
+Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production is essential to life as we know it.
+Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
 
-Over the last *** years however, the fields of biodesign and bioart have increasingly been asking similar questions not about tchnology and humanity, but technology and nature.
-In light of the increasing environmental pressures of human imposed climate change, these questions are becoming increasingly urgent.
-Technology is now being used t oreplicate and mimic natural systems.
-As Donna Haraway noted in her paper, we must thin kbeyond the human.
-\cite{haraway} The capacity of our technology has surpassed the instinctually introspective and can now mimic what is natural outside of ourselves.
-This is a critical development and harkens back to mode indigenous ways of thinking - moving and working with natural systems, rather than in opposition to.
+% With each human breath on this planet we move deeper into the anthropocene.
 
-If our relationship with technology is extending beyond the human so to must our language.
-Mori's framework is distinctly human centric.
-How then do we apply this thinking to non-human natural world?
-Is Mori's valley merely a snaller depression within a much broader valley encompassing the whole of the natural world?
-Do humans deserve a place in that broader cintext?
-To frame this question more broadly, how do we, as human beings, distinguish between what is natural and what is atrificial?
-How is this distinction shifting as our techological capacity for the artifical increases at great cost to the natural world.
+With the exception of cellular respiration, respiration in breathing organisms is a visibly mechanical process.
+Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
+Most of these devices are higly specialized and seek to mimic partifulatr aspects of respiration such as oxygen delivery, or vocal modulation.
+These works rarely mimic the actual mechanics of the components of the human body responsible for resporation favoring more abstracted and efficient mechanical systems to achieve their goal.
+Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mehcnical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
 
-Using the COVID-19 pandemic as lens and reference, this work attempts to lay a foundation for deeping inquiry into the above.
-While its focus is on a specifically human foram of respiration, respiration in broader context is omnipresent in living organisms, thus providing a jumping off point for larger scope of inquiry into the relationship between what it means to be alive and what it means to be natural.
-Something about being alive.
-\cite{zimmer}
+This work also asks us to consider the fragility of human breath in light of the COVID-19 pandemic.
+Seemingly overnight, much of the world became obsessed with breath as the tool of the virus spread.
+In the two years since, we have become incredibly attuned to the quality and sound of our own breath and what it may say about our mortality.
+This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharpy painful.
+
+% Something about being alive. \cite{zimmer}
 
@@
 
-\subsection{mechanical relevance}
+\subsection{prior work and relevance}
 % talk about the history of breathing apparatus and note that none mimic the human body
-Mechanical ventilation has been around for a long time.
+Mechanical ventilation has been around for a long time...
 Since it  seeks to provide "replacement of the respiratory muscles".
@@
 
-Two notable works in the field of kinetic sculpture provide reference, Mawa Denkye's lauching machine and another guy's breath machine thing.
+Two notable works in the field of kinetic sculpture provide reference, Maywa Denkye's lauching machine and another guy's breath machine thing.
 These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.
@@
 
-% need to figure out how to combine the two naturally
-\section{prior work}
-
-mawa deynke
-
-breath thing from bitforms
-
-homemade ventilators
+Since the focus of this piece is the noise of breath
 
@@
 
-broken down into two steps; thinking and making
-
 \subsection{system breakdown}
 
-the respiratory system is complex, so I made is simpler
+While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system.
+\cite{breathcomplexity} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
+Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchnage of oxygen as would normally occur during respiration.
 
-\subsubsection{actuation}
+\subsubsection{Actuation}
 
-how air is moved through the respiratory system
+Actuation is the system which causes air to move through the respiratory system.
+In our own bodies this is acieved by the movemne of the diaphgragm.
+In ventilators this is ofent acieved by a pump or bellows.
 
-\subsubsection{storage}
+\subsubsection{Storage}
 
-how air is stored in the respiraatory system
+Storage describes the way air is stored and contained during the act of breath.
+This is done by the lungs and is only necessary for the transfer of oxygen to the blood.
+While the lungs are arguably the most critical part of the respiratory system, they are the most difficult to simulate with accuracy and have relatively little bearing on the sound of breath in this context.
 
-\subsubsection{modulation}
+\subsubsection{Modulation}
 
-how the sound of breath is created
+Modulation is the way the sound of breath is shaped as it leaves the body.
+This is acheived by the bronchial tree, trachea, larynx, and mouth.
+After actuation has been achieved, simulation of this system ist he most relevant and interesting part of this.
 
@@
 So far a few different protptypes ahve been attempted.
-Focusing on the two main areas of inquiry, there have been auditory prototypes, exploring the effect os disembodied breath noises, as well as machnical prototypes, exploring ways to achieve a physiologically accurate breahting mechanism.
+Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect os disembodied breath noises -- as well as machnical prototypes -- exploring ways to achieve a physiologically accurate breahting mechanism.
 
@@
 \section{conclusion}
+
+The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxyneg to the blood.
+The goal of this work was to isolate the mechanical components which form the noise created by breathing.
 
@@
 \end{document}
+
+% Masahiro Mori's 1970 essay prvided the framework for describing the confusing relationship between artifically created human-like robots and, as Mori calls us, real human beings.
+\cite{mori} This framework of thought has gained significant traction over the last 50 plus years as developments in technology have made it possiblt to mimic the human body with increasing accuracy.
+This definition has also been critical in the field of computer science as we consinute ot refine mahcine learning techniques which steadily approach the capacities of the uhman brain.
+This framework describes the point at which an artificial automaton approaches a point of being eeriely similar to a human being, creating a feeling of discomfort via the uncertainty of whether it is "real".
+
+% For the lats fifty years this framework fo thought has been critical in describing technology's relationship with humanity.
+
+% Over the last *** years however, the fields of biodesign and bioart have increasingly been asking similar questions not about tchnology and humanity, but technology and nature.
+In light of the increasing environmental pressures of human imposed climate change, these questions are becoming increasingly urgent.
+Technology is now being used t oreplicate and mimic natural systems.
+As Donna Haraway noted in her paper, we must thin kbeyond the human.
+\cite{haraway} The capacity of our technology has surpassed the instinctually introspective and can now mimic what is natural outside of ourselves.
+This is a critical development and harkens back to mode indigenous ways of thinking - moving and working with natural systems, rather than in opposition to.
+
+% If our relationship with technology is extending beyond the human so to must our language.
+Mori's framework is distinctly human centric.
+How then do we apply this thinking to non-human natural world?
+Is Mori's valley merely a snaller depression within a much broader valley encompassing the whole of the natural world?
+Do humans deserve a place in that broader cintext?
+To frame this question more broadly, how do we, as human beings, distinguish between what is natural and what is atrificial?
+How is this distinction shifting as our techological capacity for the artifical increases at great cost to the natural world.
+
+%While its focus is on a specifically human foram of respiration, respiration in broader context is omnipresent in living organisms, thus providing a jumping off point for larger scope of inquiry into the relationship between what it means to be alive and what it means to be natural.
+Something about being alive.
+\cite{zimmer}

### 2021-11-29 finalproject/prototype2/script.js

@@ -1,0 +1,9 @@
+$(document).ready(function() {
+  $('.video-gallery').magnificPopup({
+  delegate: 'a', 
+  type: 'iframe',
+  gallery:{
+    enabled:true
+  }
+});
+});

### 2021-12-01 finalproject/report/ms1report.tex
new audio files and paper edits

@@
 \begin{abstract}
-the nature of breath is...
+The Nature of Breath is a kinetic sculpture exploring the sound and mechanics of human breath.
+By mimicking the mechanics of the human body this piece asks what it means for a machine to breath, drawing an uncertain line between what is considered natural and artificial.
+The Nature of Breath is also situated in response to the COVID-19 pandemic.
+In response to the virus' spread over the last two years, humans around the globe have become unusually attuned to the degrees of health or sickness that breath can indicate as well as the factors that may change one to the other.
 \end{abstract}
@@
 \keywords{Keywords}
-
+kinetic sculpture, COVID-19, biomimicry, breath
 %------------------------------
@@
 While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.
+This machine may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.
 
@@
 Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
-Most of these devices are higly specialized and seek to mimic partifulatr aspects of respiration such as oxygen delivery, or vocal modulation.
-These works rarely mimic the actual mechanics of the components of the human body responsible for resporation favoring more abstracted and efficient mechanical systems to achieve their goal.
-Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mehcnical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
+Most of these devices are highly specialized and seek to mimic particular aspects of respiration such as oxygen delivery, or vocal modulation.
+These works rarely mimic the actual mechanics of the components of the human body responsible for respiration favoring more abstracted and efficient mechanical systems to achieve their goal.
+Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mechanical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
 
 This work also asks us to consider the fragility of human breath in light of the COVID-19 pandemic.
-Seemingly overnight, much of the world became obsessed with breath as the tool of the virus spread.
+Seemingly overnight, much of the world became obsessed with breath as the toll of the virus spread.
 In the two years since, we have become incredibly attuned to the quality and sound of our own breath and what it may say about our mortality.
-This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharpy painful.
+This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharply painful.
 
@@
 
-\subsection{temporal relevance}
-% the pandemic...the soundscape of the pandemic...
-Teh COVID-19 pandemic has caused a sort of crystalization on many of the aforementioned optics, and has
+\subsection{Prior Work and Relevance}
+% talk about the history of breathing apparatus and note that none mimic the human body
+Mechanical ventilation has been around for a long time (include specific reference)...it  seeks to provide "replacement of the respiratory muscles".
+\cite{ventilatorhistory} This results in apparatuses that resemble the respiratory system only in the sense that they cause a regulated change in pressure which allows for oxygenation of the blood.
+The devices have no need, however, for recreating the process by which breath sounds are actually created.
+The field of biomimetic robotics has little use for such inquiry either -- respiration is not necessary for an electronic being and it is much easier to produce vocal noises digitally than mechanically.
 
-During the pandemic many people tried ot mkae ventilators at home. Segue...
+Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing.
+\cite{denki} \cite{lozanohemmer} These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.
 
-\subsection{prior work and relevance}
-% talk about the history of breathing apparatus and note that none mimic the human body
-Mechanical ventilation has been around for a long time...
-Since it  seeks to provide "replacement of the respiratory muscles".
-\cite{ventilatorhistory} This results in apparati that resembly the respiratory system only in the sens that they cause a regutaled chnage in pressure, but have no need for recreating htep rocess by which breaht sounds are actually created.
-Simialrly, the cield of biomemtic robotics has little use for such inquiry - respiration is not necessary for an electronic humanoid and it is much easier to produce vocal noises in a digital manner.
+% context of the pandemic ventilators
 
-Two notable works in the field of kinetic sculpture provide reference, Maywa Denkye's lauching machine and another guy's breath machine thing.
-These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.
-%\cite{deynke} \cite{otherguy}
+\section{Implementation}
 
-Since the focus of this piece is the noise of breath
-
-\section{implementation}
-
-\subsection{system breakdown}
+\subsection{System Breakdown}
 
 While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system.
-\cite{breathcomplexity} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
-Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchnage of oxygen as would normally occur during respiration.
+\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
+Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.
 
@@
 Actuation is the system which causes air to move through the respiratory system.
-In our own bodies this is acieved by the movemne of the diaphgragm.
-In ventilators this is ofent acieved by a pump or bellows.
+In our own bodies this is achieved by the movement of the diaphragm.
+In ventilators this is often achieved by a pump or bellows.
 
@@
 Modulation is the way the sound of breath is shaped as it leaves the body.
-This is acheived by the bronchial tree, trachea, larynx, and mouth.
-After actuation has been achieved, simulation of this system ist he most relevant and interesting part of this.
+This is achieved by the bronchial tree, trachea, larynx, and mouth.
+After actuation has been achieved, simulation of this system is the most relevant and interesting part of this.
 
-\subsection{making}
+\subsection{Making and Testing}
 
-So far a few different protptypes ahve been attempted.
-Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect os disembodied breath noises -- as well as machnical prototypes -- exploring ways to achieve a physiologically accurate breahting mechanism.
+So far a few different prototypes have been attempted.
+Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism.
+In addition to creating a visual tie between the biological nature of the respiratory system, I chose to closely mimic the diaphragm and trachea in this way to prevent the noise of the mechanical components from overwhelming the noise of the breath produced.
+These approaches rely on material properties of -- elasticity and rigidity -- of latex and vinyl to be functional.
 
-\subsubsection{auditory prototypes}
+\subsubsection{Auditory Prototypes}
 
@@
 
-\subsubsection{physical prototypes}
+\subsubsection{Physical Prototypes}
 
@@
 
-\subsubsection{final prototypes}
+\subsubsection{Final Prototypes}
 
@@
 
-\section{further work}
+\section{Further Work}
 
-there's always more to do...
+well, there's always more to do...
 
@@
 
-\section{conclusion}
+\section{Conclusion}
 
-The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxyneg to the blood.
+The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood.
 The goal of this work was to isolate the mechanical components which form the noise created by breathing.
 
-talk about the big distinction of artificial and natural
+bring this back to the question of the relationship between what is natural and what is alive
 
-why it's important to think about this in the context of humanity's relationship with the natural world
+here is where I can cite some of the exciting reading I've done and talk about the big distinction of artificial and natural
+
+\cite{zimmer}
+\cite{haraway}
+\cite{smith}
+
+% why it's important to think about this in the context of humanity's relationship with the natural world
 
@@
 
-\section{Bibliography}
-%The difference between a reference list and a bibliography is that in your references, you list all the sources you directly referred to in the body of your writing in numerical order, whereas a bibliography includes an alphabetical listing of all those authors and sources that you have consulted while writing your essay.
-Use the same format as for the references otherwise.
-Those using Latex will follow the usual cite command format.
-
 \section{Author Biography}
 Lauria Clarke is an engineer and artist living in New York.
-She completed an MS in Computer Engineering at Northeastern University in 2017 and is currently persuing an MFA in Design and Technology at The Parsons School of Design.
+She completed a MS in Computer Engineering at Northeastern University in 2017 and is currently pursuing an MFA in Design and Technology at The Parsons School of Design.
 She specializes in digital hardware design and kinetic sculpture and is always curious to learn more about humanity's relationship with the natural world.
@@
 % aspires to be a chocolate fish fisherman for Ben and Jerry's.
+
+\section{Note to Harpreet}
+
+Hi Harpreet,
+
+My apologies, this is quite rough.
+I plan to finish fleshing out the prior work and relevance section and then collect and polish all the writing I've done about my prototypes for the implementation section.
+I accidentally used up a lot of time writing about the uncanny valley and it's relevance in biomimetics/biodesign/biotechnology...which was far outside the scope of this paper
+
+Lauria
 

### 2021-12-01 finalproject/prototype2/script.js

@@ -1,9 +1,0 @@
-$(document).ready(function() {
-  $('.video-gallery').magnificPopup({
-  delegate: 'a', 
-  type: 'iframe',
-  gallery:{
-    enabled:true
-  }
-});
-});

### 2021-12-16 finalproject/report/ms1report.tex
report progress

@@
 
-\title{The Nature of Breath: A Biomimetic Kinetic Sculpture That Breathes}
+\title{The Nature of Breath: A Biomimetic Kinetic Sculpture that Breathes}
 \author{Lauria Clarke\\ Parsons School of Design, The New School\\ New York, USA\\ clarl404@newschool.edu\\
@@
 The Nature of Breath is a kinetic sculpture exploring the sound and mechanics of human breath.
-By mimicking the mechanics of the human body this piece asks what it means for a machine to breath, drawing an uncertain line between what is considered natural and artificial.
+By mimicking the mechanics of the human body this piece asks what it means for a machine to breath, drawing an uncertain line between what is considered natural and artificial and what is considered to be living or non-living.
 The Nature of Breath is also situated in response to the COVID-19 pandemic.
 In response to the virus' spread over the last two years, humans around the globe have become unusually attuned to the degrees of health or sickness that breath can indicate as well as the factors that may change one to the other.
+Through its contrast between the natural and artificial, this works intends to spur reflection on the pandemic as an indicator of humanity's relationship with the natural world.
 \end{abstract}
@@
 % general introduction
-
-Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive.
+% RANDOM NOTES AND OUTLINE
+% Life is tghe avoidance of entropy - an indisputable natural law of the universe, therefor its opposite, life is as well (quote from zimmer).
+%  -- is often tied up in kinetic sculpture. 
+% layering the distinctions between what is natural and artificial and what is living and non-living.    
+% 
+% Movement is a natural indicator of life which situates kinetic sculpture as an ideal lens for invetsigating this blurred boundary between the natural and artificial.    
+%   
+% kinetic sculpture often blurs the line between what is natural and artificial using movement
+% it's important that we think about this because the natural world is chnaging at an unprecidented rate
+% there is something inherently natural about movement
+% things thta move tend to be alive
+% this is a basic distinction human beings have evolved to make - between living and nonliving
+% through movement there is a deep relationship bewteen what is natural and what is or has been alive
+% this work uses the movement of breath to explore this relationship
+% CONCLUSION:
+% Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive.
 While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.
 This machine may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.
-
-% Respiration is an indicator of life at all scales - from single celled organisms to human beings.
-Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production is essential to life as we know it.
-Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
-
 % With each human breath on this planet we move deeper into the anthropocene.
-
-With the exception of cellular respiration, respiration in breathing organisms is a visibly mechanical process.
-Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
+% Respiration in breathing organisms is a visibly mechanical process. 
+% Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
 Most of these devices are highly specialized and seek to mimic particular aspects of respiration such as oxygen delivery, or vocal modulation.
@@
 Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mechanical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
+
+% art history reference https://thekeep.eiu.edu/cgi/viewcontent.cgi?article=4843&context=theses
+
+Art has been used as a medium to reflect and celebrate the beauty of the natural world for a long time.
+As the machine age took hold of the public's imagination and facility with mechanical devices became wide-spread, artist began to turn to medium of movement to reflect the natural world CITE.
+The birth of modern computing has made small, easy to use, and low voltage electronics increasingly accessible, allowing a new crop of artists to create even wider range of movement for mirrorig the complex beauty of movement in the natural world.
+
+The process of re-creating natural movement through mechanical, man-made means allows the artist to highlight and isolate natural movement outside of its conventional context in nature.
+Mimicing the movement of a field of wheat in the wind, for example, draws our attention from the awe-inspiring scale of such a landscape to focus the magical quality of movement produced by a single stalk operating within a group.
+This technique is incredbily powerful, providing new context for the beauty of movement often overlooked in everyday life.
+While isolating such an innate quality of the natural world is itself magical, those looking a little closer cannot help but notice the stark contrast between such a natural movement and the artificial way in which it may be achieved.
+This blurred distinction between the natural and the artificial is a hallmark characteristic of biomimetic movement from kinetic sculpture to engineering and beyond.
+
+Our present understanding and knowledge of the natural world and it boundaries is changing at an unprecidented rate due to human caused environmental change.
+The question of what is natural and what has been altered, enhanced, or created by humankind -- the artificial -- is therefore increasingly urgent and difficult to answer.
+Through the lens of this question it becomes increasingly apparent that creating a solid boundary between humankind and the rest of the natural world is a mindset which cannot be sustained indefinitely.
+As Jedediah Purdy remarks in his book, "the natural and the artificial have merged at every scale...If Nature were a place, we could not find it.
+If Nature were a state of mind, we could not attain it.
+We are something else, and so is the world." CITE
+
+The question of what it means to be alive is unavoidable when considering the relationship between movment and what is natural.
+Movement is a key indicator of life and the ability to distinguish between the living and nonliving is a deeply rooted capacity of the human brain.
+Psychologists and Neurologists have long been aware that "the distinction between living and nonliving things forms a natural place," to divide the funtionlities and capacities of the mind CITE.
+Kinetic sculpture layers these two distinctions -- between the natural and the arificial and the living an non-living -- in an open-ended way which scienetific disciplines are rarely allowed.
+
+Perhaps the most natural indicator of life is the movement of breathing.
+Respiration is an indicator of life at all scales - from single celled organisms to human beings.
+Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it.
+Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
+This mechanism may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.
+There is no question that this breathing work does not actually posses life.
 
@@
 This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharply painful.
-
 % Something about being alive. \cite{zimmer}
 
-\subsection{Prior Work and Relevance}
+\subsection{Prior Work}
 % talk about the history of breathing apparatus and note that none mimic the human body
-Mechanical ventilation has been around for a long time (include specific reference)...it  seeks to provide "replacement of the respiratory muscles".
+Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing.
+\cite{denki} \cite{lozanohemmer} Denki's work is arguable the more mechanically sophisticated of the two and most relevant in the discussion of breathing noises.
+WAHHA GO GO uses a bellows to force air through a larynx like opening which expands and contracts producing a laughing noise.
+The translation of motion through this piece is done using extremely precise and well made metal parts.
+This means that little noise -- aside from a pleasant whirring sound -- is generated by the mechanism itself focusing all of the viewers' attention on the eerie laughter it produces.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/vcb_madrid_2014.png}
+\caption{Vicious Circular Breathing by Rafael Lozano-Hemmer}
+\end{figure}
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/whagogo.png}
+\caption{WAHA GO GO by Maywa Denki}
+\end{figure}
+
+In contrast, Vicious Circular Breathing relies on the sharp sound of crinking paper bags to indicate the movement of breath.
+This behemoth machine composed of gigantic bellows, yards of tubing, and a huge array of relays makes a lot of noise.
+Lozano-Hemmer's interest in breath is less about its bodily mechanics and more about the impacts and (un)sustainability of human breathing.
+Users are invited to add their breath to the closed loop system which eventually approaches a level of CO$_2$ saturation, rendering the piece unusable.
+Ultimately this sends a darker message about human breath.
+
+These two pieces provoke thought about breath, but still use simple bellow based mechanisms to produce airflow.
+Bellows have been used for a long time in mechanical ventilation.
+Mechanical ventilation, used in medicine since A LONG TIME seeks to provide "replacement of the respiratory muscles".
 \cite{ventilatorhistory} This results in apparatuses that resemble the respiratory system only in the sense that they cause a regulated change in pressure which allows for oxygenation of the blood.
@@
 
-Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing.
-\cite{denki} \cite{lozanohemmer} These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.
-
-% context of the pandemic ventilators
+Finally, it would be an oversight not to mention the hundreds of engineers and makers who devoted much time and though to the development of opensource emergency ventilators during the beginning and worst months of the pandemic.
+Very few of these devices were ever used on patients, but the collective thought and conversation surrounding the mechanics of breath cannot be discounted as useless, if only as a poignant reminder of the body's mystery and vulnerability.
 
@@
 
-While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system.
+While not the most complex respiratory system on the planet, the human respiratory system is a very complicated biological system.
 \cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
 Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/respiratorysystem.png}
+\caption{Diagram of the human respiratory system}
+\end{figure}
 
@@
 Modulation is the way the sound of breath is shaped as it leaves the body.
-This is achieved by the bronchial tree, trachea, larynx, and mouth.
-After actuation has been achieved, simulation of this system is the most relevant and interesting part of this.
+This is achieved by the bronchial tree, trachea, larynx, pharynx, and mouth CITE.
+After actuation has been achieved, simulation of this system is the most relevant and interesting part of this project.
 
@@
 
-\subsubsection{Auditory Prototypes}
+\subsubsection{Auditory Trials}
 
-bucket experiement
+The obective of these prototypes were to understand the reaction observers have ot disembodied breath noises.
+Disembodied in this case means that the noise is not coming from a source which appeasr to be alive.
+Some of this testing was done in a classroom with an audience who knew what was happening, and some was done on the streets of New York City with passesby.
+Both achieved surprisingly similar results.
 
-p5 thing
+To test the effect of disembodied breath on strangers, a small speaker was placed inside of two five gallon plastic buckets.
+An old cellphone was pre-loaded with a three minute soundeffect track of different types of breathing noises.
+These cycled through breath which sounded like someone sleeping to someone wheezing from exertion.
+In one bucket, a hole was drilled and a large diameter latex tube was inserted FIG.
+This provided a way to direct the sound eminating from the speaker and added a level of visual confusion to the assembly.
+The assembly was then placed near a crosswalk with a high level of traffic around 8 PM ET on the day of Halloween.
+This location was chosen specifically because it was easy to observe discreetly and had a wide variety of foot traffic.
 
-\subsubsection{Physical Prototypes}
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/buckettest.png}
+\caption{Passesby observing a breathing bucket.}
+\end{figure}
 
-trachea
+Due to the roadnoise of the location we only observed a handful of passersby alter their activity because of the bucket.
+It's possible many more heard the noise and simply ignored it.
+Those who did notice the noises were visibly consused or disconcerted and usually took a moment to look inside the bucket.
+It was clear base on this action that they were expecting something to be inside the bucket.
+The most interesting interaction happened while a young woman was waiting at the crosswalk with a group of friends.
+She happened to be near the edge of the group closest to the bucket and was clearly confused yet intriuged by the noise as she kept looking furtively over her shoulder at the bucket.
+It was also clear that she felt some uncertainty or concern because we observed that she didn’t immediately mention it to her companions.
 
-diaphragm
+The second investigation of disembodied breath happened in a classroom setting with a prrepared audience.
+First, normal breath noises were played.
+Then, the noises of someone coughing and wheezing were overlaid at a much louder volume.
+Almost immideately, the audience was extremely uncomfortable with the "sick" noises.
+A few began to cover their ears.
+Due to the level of discomfort visible around the room we decided to end the experiment.
+A smaller version of this experiment was repeated at lower volume with an audience of only three people with similar results.
+There was immediate, visible concern among the audience when the "sick" breathing began to play.
 
-lever adjustment
+\subsubsection{Physical Trials}
 
-\subsubsection{Final Prototypes}
+The objective of these second category of prototypes was to investigate way of generating breathing noises mechanically.
+Using the systems breakdown discussed above, we attemted to create mechanisms for both modulation and actuation of breath.
+We first aproached the diaphragm since acuation was a necessary step to fully explore modulation.
 
-the final result was a thing that nearly works and combines many of the elements from above
+Using the five gallon buckets earlier, we cut out the center of the the lid of one bucket and used it to secure a large piece of 25 mil natural latex over the opening.
+We then inserted the same large piece of latex tubing into a hole at the other end of the bucket FIG.
+By depressing the latex air was forced out pf the tube.
+Pinching the tube at different heights and changing the cadence of depressing the latex created a pretty wide range of breath-like noises.
+
+%\begin{figure}[h]
+%\includegraphics[width=0.5\textwidth]{images/diaphragm.png}
+%\caption{Diaphragm and rudimentary trachea using five gallon bucket.}
+%\end{figure}
+
+This assembly was then mounted on a frame created using t-slotted aluminum.
+A lever was also installed under the diaphgram FIG.
+The end of the leve had a long piece of metal wrapped in a soft rag which was roughly centered on the diaphragm.
+By depressing the lever with our feet the end of thie lever extended into the diaphragm activating it while we used our hands to createvariation in the noise of the airflow through the tube.
+This was sort of a "playable" breath maker.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/playablebreath1.png}
+\caption{Previous assembly installed with lever.}
+\end{figure}
+
+After measuing the force required to depress the lever fully -- 100 N -- we determiend that creating an electronic system to activate the diaphragm would be costly and complicated given the scope of time.
+A few attempts at creating mechnical advantage for the lever were attempted, but none had successful results.
+It was very difficult to generate the required amount of force without putting a significant amount of weight on the lever itself.
+We ultimately decided that, due to the expense of high-powered and quiet motors, it made more sense to solve this problem after the core interaction of the piece had been refined.
+
+To approach modulation we decided to rely, again, on the mechnics of the human body.
+The trachea contains C shaped bands of caralidge with a section smooth muscle on the open side.
+When this muscle contracts the C shape is compressed changing how air flows through the trachea.
+The trachea is about one inch in diamameter CITE.
+We found that a two inch section of $1/8$ inch vinyl tubing had an internal radius of about one inch when bent in a C shape which reduced to less than a quarter of an inch when the two ends of teh C were pulled together.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/trachea.png}
+\caption{Cross section of the trachea and esophagus.}
+\end{figure}
+
+We embedded four sections of tubing between two sheets of fourteen mil natural latex.
+We then bent this into a curved tub-like shape and enclosed the open end of the C FIG.
+Once this basic shape was complete we threaded each piece of tubing with monofillament and anchored the assembly between two dowels.
+By pulling on the monofillament is was possible to contract each ring within the tube individually FIG.
+This result was highly successful and it was possible to contract the aperature to varrying degrees.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.22\textwidth]{images/46.JPG}
+\includegraphics[width=0.22\textwidth]{images/52.JPG}
+\caption{Sequence of contracting each ring then releasing all of them slowly}
+\end{figure}
+
+While this result was not intergrated into the modulation prototype, it was by far the most naturalistic movement produced during the investigation.
+In comparison to the latex tubing used to minic thre trachea earlier, this mechanical latex assembly has much thinner walls.
+Overall, we belive this will produce more accurate breath noises when combined with the actuation mechanism
+
+\subsubsection{Combined Trial}
+
+Based on the results of the trials discussed above, we produced a final version combining the successful auditory and physcial components.
+
+We first created a digital breath generator in order to create a more nuanced interaction between the viewer and the breathing noises.
+This was done using p5.js and allowed us to change the cadence, pitch, and volume of the breathing noise.
+A simple infra read distance sensor was then connected to the generator.
+As the distance between the viewer and the work decreased the breathing increased in intensity.
+The breath noise was amplified using a class D audio amplifier and played through speakers mounted on the front of the piece.
+
+%\begin{figure}[h]
+%\includegraphics[width=0.5\textwidth]{images/digitalbreath.png}
+%\caption{Digital breath generator.}
+%\end{figure}
+
+The mechanical form of the work stayed the same in this trial, however we removed the five gallon bucket in favor of a clear vessle which allowed the viewer to see the diaphragm moving.
+We also added an abstracted bronchial tree inside the vessel to suggest lungs.
+In combination with the added electronics this had a much more polished effect.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/finaltrial.JPG}
+\caption{Final form with infra-red distance sensor and digital bretah generator.}
+\end{figure}
+
+This trial was much more effective in two regards.
+The auditory component of the interaction was much more natural and made it easier to grab the viewer's attention.
+The visual appearance of the piece was also more intersting and made it easier to draw the connection between the object and the bretahing noises produced.
+
+% First, the interaction between the viewer and the breathing noises was much more intuitive.
+In prior user interviews we found that having the intensity of the breathing incerase in the presence of the viewer was more intuitive than the intensity of breath decreasing.
+This proved to be true as the change in intensity of breathing piqued viewers interest, causing them to approach the work.
+When they were vey close, the breathing reached its highest level of intensity and many viewers noted that they felt somewhat uncomfortable to be in such close proximity to the piece.
+In comparison, some users mentioned that at far away distances -- producing slow and relaxed breathing noises -- they found their own breath falling into a relaxing rythym and were enjoyed the meditative quatily produced.
+Overall, the auditory interaction consistently caused viewers to consider their own breath.
+
+\section{Discussion}
+
+% FIXME
+
+This result was interesting because it illustrated just how sensitive the group had be come to the noise of respiratory distress.
+While driect comparison is not possible, we wonder if this experiment would have had the same effect before the COVID-19 pandemic.
+It was also clear from the repitition of this trial that volume played a role in the levl of discomfort.
+Loud bretahing noises inherently feel closer and more intimate.
+Additionally, we found that the cadence and pitch of breathing noises had a lot of bearing on the emotion of the observer.
+A fast cadence of slighly higher pitch at higher volume was more noticible and produced a higher degree of discomfort than a slower, softer breathing noise.
+
+Initial research into the effects of disembodied breath revealed that the intensity and volume of the breathing noise has a large impac on the comfort level of the viewer.
+Louder, more intense bretahing noises sounded more intimate and closer ultimately making the user feel more uncomfortable in their presence.
+
+We used this result to develop the digital bretah generator which could be tuned to change the intensity and volume of bretahing noises based on user proximity.
+This created a much more intuitive interaction between the viewer and the breathing noise.
+In prior user interviews we found that having the intensity of the breathing incerase in the presence of the viewer was more natural than the intensity of breath decreasing.
+This proved to be true as the change in intensity of breathing piqued viewers interest, causing them to approach the work.
+When they were vey close, the breathing reached its highest level of intensity and many viewers noted that they felt somewhat uncomfortable to be in such close proximity to the piece.
+In comparison, some viewers mentioned that at far away distances -- producing slow and relaxed breathing noises -- they found their own breath falling into a relaxing rythym and were enjoyed the meditative quatily produced.
+Overall, the auditory interaction consistently caused viewers to consider their own breath.
+
+While the variation in breath noises that could be produced using the initial trachea and diaphragm mechanism was very satisfactiry, the look and feel of the mechanism had little relationship to the human body.
+Viewers noted that it was hard to relate the noised produced to the object itself because of this.
+By making the movemnet of the diaphragm visible and creating an internal structure reminscent of the broncial tree we were able to draw viewers attention to the intended functionality of the work without the additional context of movement or sound.
+Upon seeing the bronchial structure many viewers immidiately concluded that the work had to do with the lungs.
+This static visual cue really helped to inforce the intent of the interaction.
+
+The bronchial tree added an immediate visual indicator of the work's functionality while the addition of the clear chamber allowed the viewer to guess at the diaphragm's funtionality.
+During user testing when the lever mechanism was activated viewers were extremely cusious about the movement of the diaphragm.
+This simple motion was the most intuitive and seemed to produce the most natural reference to breath.
+Based on the sound eminating from the broncial tree and trachea it was very clear that the motion of diaphragm was causing air to be forced out of the vessel.
 
@@
 
-well, there's always more to do...
+% emotion of breath
+% mechanical complexity
+% context
 
-the boris vian reference
+While the combination of mechanical breath and user interaction are core components of this exploration, they are not strictly necessary to continue exploring the emotion which can be conveyed through breath.
+The auditory trials in this investigation only scratched the surface of the nuanced emotion that breath can signify.
+I am curious to use this work as a platform to explore the narrative qualities of breath through viewer interaction.
+What kind of dialouge can the work spark between itself and the viewer or between multiple viewers?
+What types of stories can the viewer inuit between themself and the work?
+These questions ultimately ask what range of emotion can we convey through the act of breathing.
+
+Expanding on the mechanical techniques developed here, I hope to create a fully automated, mechanical version of this work.
+Combining the interatcive qualities of the digital breath generator with the visual qualities of the mechanical trials will ultimately tie the core interaction and the kinetic intent of this work together.
+This investigation intended to pinpoint the relationship between what is natural and what is alive through the medium of movement.
+User testing revealed that the kinetic potential of this work was the most exciting part.
+However, this work is not yet living since it does not intitiate the visble act of breathing.
+
+By replacing the lever mechanism with a crank shaft -- a technique for translating circula rmotion into linear motion -- I hope to have fine grained control over the movement of the diaphragm and thereby the actuation of air through the work.
+This will then allow for further development and testing of the mechanical trachea.
+The modulation of breath does not end with the trachea however, and I hope to continue exploring the ways in which this sound can be manipulated.
+Perhaps the developemnt of a mechanical pharynx, or larynx
+
+Finally, I am curious to investigate how the contex of this work can be expanded beyond a traditional gallery space.
+One of the original questions this work intended to address, alebit obliquely, was how breath can be extended as a metaphor for humanity's relationship with the natural world.
+I believe it will be difficult to answer this question more specifically within the constraints of a gallery.
+Putting this piece in conversation with the natural world itself would expand the relationship of its breath beyond the viewer to the environment it inhabits.
 
@@
 
+Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive.
+While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.
+This machine may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.
+
 The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood.
-The goal of this work was to isolate the mechanical components which form the noise created by breathing.
+The mechanical goal of this work was to isolate the  which form the noise created by breathing.
+
+James Prosek
+What can art provide in this time when our earth is indeed looking vulnerable to rapid and regrettable change induced by human influence?
+
+Mark Dion
+Your question answers itself - the job of the artist-naturalist is precisely to interrogate the question of nature when it is subject to rapid change and its very definition is evolving in unprecedented ways.
 
@@
 here is where I can cite some of the exciting reading I've done and talk about the big distinction of artificial and natural
+
+The root of my curiosity in addressing the boundary between what is natural and artificial is a desire to understand how creative practitioners can use the uncertainty surrouding thie distinction to generate reflection and consideration of humanity's place in the natural world.
+
+To conceptualize the impact of human activity on this planet is practically impossible, yet we find ourselves at a juncture where this definition is at the heart of nearly every social and political issue.
+If Art is a trascendent mechanism for collapsing the scales of Time and space we need it now more than ever to make sense of this recursively enormous question.
+
+% CONCLUSION:
+% With each human breath on this planet we move deeper into the anthropocene.
+% Respiration in breathing organisms is a visibly mechanical process. 
+% Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
+Most of these devices are highly specialized and seek to mimic particular aspects of respiration such as oxygen delivery, or vocal modulation.
+These works rarely mimic the actual mechanics of the components of the human body responsible for respiration favoring more abstracted and efficient mechanical systems to achieve their goal.
+Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mechanical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
 
@@
 Lauria Clarke is an engineer and artist living in New York.
-She completed a MS in Computer Engineering at Northeastern University in 2017 and is currently pursuing an MFA in Design and Technology at The Parsons School of Design.
 She specializes in digital hardware design and kinetic sculpture and is always curious to learn more about humanity's relationship with the natural world.
 
+% She completed a MS in Computer Engineering at Northeastern University in 2017 and is currently pursuing an MFA in Design and Technology at The Parsons School of Design.
+
 % aspires to be a chocolate fish fisherman for Ben and Jerry's.
-
-\section{Note to Harpreet}
-
-Hi Harpreet,
-
-My apologies, this is quite rough.
-I plan to finish fleshing out the prior work and relevance section and then collect and polish all the writing I've done about my prototypes for the implementation section.
-I accidentally used up a lot of time writing about the uncanny valley and it's relevance in biomimetics/biodesign/biotechnology...which was far outside the scope of this paper
-
-Lauria
 

### 2021-12-16 finalproject/report/ms1report.tex
nearly done with paper

--- 2021-12-01 finalproject/report/ms1report.tex
@@
+%TODO
+% add figure references
+% finish conclusion
+% add diaphragm and bronchial tree closeups
+% review reference ordering
+
 \documentclass[letterpaper]{article}
@@
 \usepackage[numbers]{natbib}
+\raggedbottom
+
 \pdfinfo{/Title (The Nature of Breath) /Author (Lauria Clarke)}
 
-\title{The Nature of Breath: A Biomimetic Kinetic Sculpture That Breathes}
-\author{Lauria Clarke\\ Parsons School of Design, The New School\\ New York, USA\\ clarl404@newschool.edu\\
-\newline
-\newline
-}
+\title{The Nature of Breath: A Biomimetic Kinetic Sculpture that Breathes}
+\author{Lauria Clarke\\ Parsons School of Design, The New School\\ New York, USA\\ clarl404@newschool.edu\\}
 \setcounter{secnumdepth}{0}
@@
 
+% art history reference https://thekeep.eiu.edu/cgi/viewcontent.cgi?article=4843&context=theses
+
+% Art has been used as a medium to reflect and celebrate the beauty of the natural world for a long time. [need sentence here tying to the next paragraph]
+%As the machine age took hold of the public's imagination and facility with mechanical devices became wide-spread, artist began to turn to medium of movement to reflect the natural world CITE.
+The birth of modern computing has made small, easy to use, and low voltage electronics increasingly accessible, allowing a new crop of artists to create even wider range of movement for mirrorig the complex beauty of movement in the natural world.
+
+Kinetic sculpture can be an incredibly powerful technique for re-contextualizing the beauty of movement often overlooked in everyday life.
+The process of recreating natural movement through mechanical, man-made means allows artists to highlight and isolate natural movement outside of its conventional context in nature.
+Mimicking the movement of a field of wheat in the wind, for example, draws our attention from the awe-inspiring scale of such a landscape to focus on the magical quality of movement produced by a single stalk operating within a group.
+While isolating such an innate quality of the natural world is itself magical, those looking a little closer cannot help but notice the stark contrast between such a natural movement and the artificial way in which it may be achieved.
+This blurred distinction between the natural and the artificial is a hallmark characteristic of biomimetic movement from kinetic sculpture to engineering and beyond.
+
+Our present understanding and knowledge of the natural world and it boundaries is changing at an unprecedented rate due to human caused environmental change.
+The question of what is natural and what has been altered, enhanced, or created by humankind -- the artificial -- is therefore increasingly urgent and difficult to answer.
+Through the lens of this question it becomes clear that creating a solid boundary between humankind and the rest of the natural world is a mindset which cannot be sustained indefinitely.\cite{smith} As Jedediah Purdy notes, ``the natural and the artificial have merged at every scale...If Nature were a place, we could not find it.
+If Nature were a state of mind, we could not attain it.
+We are something else, and so is the world.''\cite{purdy}
+
+The question of what it means to be alive is unavoidable when considering the relationship between movement and what is natural.
+Movement is a key indicator of life and the ability to distinguish between the living and non-living is a deeply rooted capacity of the human brain.
+Psychologists and Neurologists have long been aware that ``the distinction between living and non-living things forms a natural place,'' to divide the functionalities and capacities of the mind.\cite{narine} Kinetic sculpture layers these two distinctions -- between the natural and the artificial and the living and non-living -- in an open-ended way which scientific disciplines are rarely allowed.
+
+Perhaps the most natural indicator of life is the movement of breathing.
+Respiration is an indicator of life at all scales - from single celled organisms to human beings.
+Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it.
+Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
+This mechanism may breathe in a natural way, but its appearance clearly identifies it as something which is purely artificial.
+There is no question that this breathing work does not actually posses life.
+
 This work also asks us to consider the fragility of human breath in light of the COVID-19 pandemic.
@@
 In the two years since, we have become incredibly attuned to the quality and sound of our own breath and what it may say about our mortality.
-This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead breath to be easy or sharply painful.
-
+This work seeks to probe these newfound sensitivities, forcing us to question which actions may lead the act of breathing to be easy or sharply painful.
 % Something about being alive. \cite{zimmer}
 
-\subsection{Prior Work and Relevance}
+\subsection{Prior Work}
 % talk about the history of breathing apparatus and note that none mimic the human body
-Mechanical ventilation has been around for a long time (include specific reference)...it  seeks to provide "replacement of the respiratory muscles".
-\cite{ventilatorhistory} This results in apparatuses that resemble the respiratory system only in the sense that they cause a regulated change in pressure which allows for oxygenation of the blood.
-The devices have no need, however, for recreating the process by which breath sounds are actually created.
+Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing.\cite{denki}\cite{lozanohemmer}
+
+Denki's work is arguably the more mechanically sophisticated of the two and most relevant in the discussion of breathing noises.
+WAHHA GO GO uses a bellows to force air through a larynx like opening which expands and contracts producing a laughing noise.
+The translation of motion through this piece is done using extremely precise and well made metal parts.
+This means that little noise -- aside from a pleasant whirring sound -- is generated by the mechanism itself focusing all of the viewers' attention on the eerie laughter it produces.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/whagogo.png}
+\caption{WAHA GO GO by Maywa Denki}
+\label{fig:denki}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/vcb_madrid_2014.jpg}
+\caption{Vicious Circular Breathing by Rafael Lozano-Hemmer}
+\label{fig:hemmer}
+\end{figure}
+
+In contrast, Vicious Circular Breathing relies on the sharp sound of crinkling paper bags to indicate the movement of breath.
+This behemoth machine composed of gigantic bellows, yards of tubing, and a huge array of relays makes a lot of noise.
+Lozano-Hemmer's interest in breath is less about its bodily mechanics and more about the impacts and (un)sustainability of human breathing.
+Users are invited to add their breath to the closed loop system which eventually approaches a level of CO$_2$ saturation, rendering the piece unusable.
+Ultimately this sends a darker message about human breath.
+
+These two pieces provoke thought about breath, but still use simple bellow based mechanisms to produce airflow.
+Bellows have been used for a long time in mechanical ventilation.
+Mechanical ventilation, as used in medicine since the early 1900s, seeks to provide ``replacement of the respiratory muscles.''\cite{ventilatorhistory} This results in apparatuses that resemble the respiratory system only in the sense that they cause a regulated change in pressure which allows for oxygenation of the blood.
+These devices have no need for recreating the process by which breath sounds are actually produced.
 The field of biomimetic robotics has little use for such inquiry either -- respiration is not necessary for an electronic being and it is much easier to produce vocal noises digitally than mechanically.
 
-Two notable works in the field of kinetic sculpture provide some reference and context for this work, Maywa Denki's WAHHA GO GO and Rafael Lozano-Hemmer's Vicious Circular Breathing.
-\cite{denki} \cite{lozanohemmer} These two pieces provoke thought about breath, but still use simple bellow bases mechanisms to produce airflow.
-
-% context of the pandemic ventilators
+Finally, it would be an oversight not to mention the hundreds of engineers and makers who devoted much time and though to the development of OpenSource emergency ventilators during the beginning and worst months of the pandemic.\cite{read} Very few of these devices were ever used on patients, but the collective thought and conversation surrounding the mechanics of breath cannot be discounted as useless, if only as a poignant reminder of the body's mystery and vulnerability.
 
@@
 
+Before beginning construction of the breathing machine I made a cursory, but interesting investigation into some of the techniques used for creating pressure differentials of air in industrial processes.
+These were expensive, noisy, and ultimately lay outside the scope of this discussion.
+The intent of this work was to draw mechanical inspiration from the human body, so I began my inquiry there.
+
 \subsection{System Breakdown}
 
-While the avian respiratory system is, in fact, more complex, the human respiratory system is a very complicated biological system.
-\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
+While not the most complex respiratory system on the planet, the human respiratory system is very complicated.\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
 Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/respiratorysystem.png}
+\caption{Diagram of the human respiratory system.}
+\end{figure}
 
@@
 
-Actuation is the system which causes air to move through the respiratory system.
+Actuation is the process which causes air to move through the respiratory system.
 In our own bodies this is achieved by the movement of the diaphragm.
@@
 Modulation is the way the sound of breath is shaped as it leaves the body.
-This is achieved by the bronchial tree, trachea, larynx, and mouth.
-After actuation has been achieved, simulation of this system is the most relevant and interesting part of this.
+This is achieved by the bronchial tree, trachea, larynx, pharynx, and mouth.\cite{sarkar} After actuation has been achieved, simulation of this system is the most relevant and interesting part of this project.
 
@@
 
-So far a few different prototypes have been attempted.
-Focusing on the two main areas of inquiry, there have been auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism.
-In addition to creating a visual tie between the biological nature of the respiratory system, I chose to closely mimic the diaphragm and trachea in this way to prevent the noise of the mechanical components from overwhelming the noise of the breath produced.
-These approaches rely on material properties of -- elasticity and rigidity -- of latex and vinyl to be functional.
+Focusing on the two main areas of inquiry, the making and testing has been divided into been auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism.
+In addition to creating a visual tie between the biological nature of the respiratory system, I chose to closely mimic the diaphragm and trachea to prevent the noise of the mechanical components from overwhelming the noise of the breath produced.
+This approach relies on the material properties of latex and vinyl -- elasticity and rigidity -- to be functional.
 
-\subsubsection{Auditory Prototypes}
+\subsubsection{Auditory Trials}
 
-bucket experiement
+The objective of these prototypes was to understand the reaction observers have to disembodied breath noises.
+Disembodied in this case means that the noise is not coming from a source which appears to be alive.
+Some of this testing was done in a classroom with an audience who knew what was happening, and some was done on the streets of New York City with passersby.
+Both achieved surprisingly similar results.
 
-p5 thing
+To test the effect of disembodied breath on strangers, a small speaker was placed inside of two five gallon plastic buckets.
+An old cellphone was pre-loaded with a three minute sound effect track of different types of breathing noises.
+These cycled through breath which sounded like someone sleeping to someone wheezing from exertion.
+In one bucket, a hole was drilled and a large diameter latex tube was inserted.
+This provided a way to direct the sound emanating from the speaker and added obfuscated the function of the assembly.
+The assembly was then placed near a crosswalk with a high level of foot traffic around 8 PM ET on the day of Halloween.
+This location was chosen specifically because it was easy to observe discreetly and had a wide variety of pedestrians.
 
-\subsubsection{Physical Prototypes}
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/buckettest.png}
+\caption{Passesby observing a breathing bucket.}
+\end{figure}
 
-trachea
+Due to the road noise of the location I only observed a handful of passersby alter their activity because of the bucket.
+It's possible many more heard the noise and simply ignored it.
+Those who did notice the noises were visibly confused or disconcerted and usually took a moment to look inside the bucket.
+It was clear based on this action that they were expecting something to be inside the bucket.
+The most interesting interaction happened while a young woman was waiting at the crosswalk with a group of friends.
+She happened to be near the edge of the group closest to the bucket and was clearly confused yet intrigued by the noise, as she kept looking furtively over her shoulder at the bucket.
+It was also clear that she felt some uncertainty or concern about the interaction because I observed that she didn’t immediately mention it to her companions.
 
-diaphragm
+The second investigation of disembodied breath happened in a classroom setting with a prepared audience.
+First, normal breath noises were played.
+Then, the noises of someone coughing and wheezing were overlaid at a much louder volume.
+Almost immediately, the audience was extremely uncomfortable with the ``sick'' noises.
+A few began to cover their ears.
+Due to the level of discomfort visible around the room I decided to end the experiment.
+A smaller version of this experiment was repeated at lower volume with an audience of only three people with similar results.
+There was immediate, visible concern among the audience when the ``sick'' breathing began to play.
 
-lever adjustment
+\subsubsection{Physical Trials}
 
-\subsubsection{Final Prototypes}
+The objective of this second category of trials was to investigate ways of generating breathing noises mechanically.
+Using the systems breakdown discussed above, I attempted to create mechanisms for both modulation and actuation of breath.
+I first approached the diaphragm since actuation was a necessary step to fully explore modulation.
 
-the final result was a thing that nearly works and combines many of the elements from above
+Using the five gallon buckets earlier, I cut out the center of the lid of one bucket and used it to secure a large piece of 25 mil natural latex over the opening.
+I then inserted the same large piece of latex tubing into a hole at the other end of the bucket.
+By depressing the latex air was forced out of the tube.
+Pinching the tube at different heights and changing the cadence of the diaphragm movement created a pretty wide range of breath-like noises.
 
+%\begin{figure}[h]
+%\includegraphics[width=0.5\textwidth]{images/diaphragm.png}
+%\caption{Diaphragm and rudimentary trachea using five gallon bucket.}
+%\end{figure}
+
+This assembly was then mounted on a frame created using t-slotted aluminum.
+A lever was also installed under the diaphragm.
+The end of the lever had a long piece of metal wrapped in a soft rag which was roughly centered on the diaphragm.
+By depressing the lever with my foot the end of this lever extended into the diaphragm activating it while I used my hands to create variation in the noise of the airflow through the tube.
+This was sort of a ``playable'' breath maker.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/playablebreath1.png}
+\caption{Previous assembly installed with lever.}
+\end{figure}
+
+After measuring the force required to depress the lever fully -- 100 N -- I determined that creating an electronic system to activate the diaphragm would be costly and complicated given the scope of time.
+A few attempts at creating mechanical advantage for the lever were attempted, but none had successful results.
+It was very difficult to generate the required amount of force without putting a significant amount of weight on the lever itself.
+I ultimately decided that, due to the expense of high-powered and quiet motors, it made more sense to solve this problem after the core interaction of the piece had been refined.
+
+To approach modulation I decided to rely, again, on the mechanics of the human body.
+The trachea contains C shaped bands of cartridge with a section smooth muscle on the open side.
+When this muscle contracts the C shape is compressed changing how air flows through the trachea.
+The trachea is about one inch in diameter.\cite{furlow} I found that a two inch section of $1/8$ inch vinyl tubing had an internal radius of about one inch when bent in a C shape which reduced to less than a quarter of an inch when the two ends of the C were pulled together.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/trachea.png}
+\caption{Cross section of the trachea and esophagus.}
+\end{figure}
+
+I embedded four sections of this tubing between two sheets of fourteen mil natural latex.
+I then bent this into a curved tube-like shape and enclosed the open end of the C FIG.
+Once this basic shape was complete I threaded each piece of tubing with monofilament and anchored the assembly between two dowels.
+By pulling on the monofilament is was possible to contract each ring within the tube individually.
+This result was highly successful and it was possible to contract the aperture to varying degrees.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.22\textwidth]{images/46.JPG}
+\includegraphics[width=0.22\textwidth]{images/52.JPG}
+\caption{Contracted and relaxed trachea.}
+\end{figure}
+
+While this result was not integrated into the modulation prototype, it was by far the most naturalistic movement produced during the investigation.
+In comparison to the latex tubing used to mimic the trachea earlier, this mechanical latex assembly has much thinner walls.
+Overall, I believe this will produce more accurate breath noises when combined with the actuation mechanism
+
+\subsubsection{Combined Trial}
+
+Based on the results of the trials discussed above, I produced a final version combining the successful auditory and physical components.
+
+I first created a digital breath generator in order to create a more nuanced interaction between the viewer and the breathing noises.
+This was done using P5.js and allowed me to change the cadence, pitch, and volume of the breathing noise.
+A simple infra-red distance sensor was then connected to the generator.
+As the distance between the viewer and the work decreased the breathing increased in intensity.
+The breath noise was amplified using a class D audio amplifier and played through speakers mounted on the front of the piece.
+
+%\begin{figure}[h]
+%\includegraphics[width=0.5\textwidth]{images/digitalbreath.png}
+%\caption{Digital breath generator.}
+%\end{figure}
+
+The mechanical form of the work stayed the same in this trial, however I removed the five gallon bucket in favor of a clear vessel which allowed the viewer to see the diaphragm moving.
+I also added an abstracted bronchial tree inside the vessel to suggest lungs.
+In combination with the added electronics this had a much more polished effect.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/finaltrial.JPG}
+\caption{Final form with infra-red distance sensor and digital breath generator.}
+\end{figure}
+
+This trial was much more effective.
+The auditory component of the interaction was much more natural and made it easier to grab the viewer's attention.
+The visual appearance of the piece was also more interesting, making it easier to draw the connection between the object and the breathing noises produced.
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/closeupdiaphragm.JPG}
+\caption{Diaphragm being activated by lever.}
+  \label{fig:denki}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=0.5\textwidth]{images/closeupbronchialtree.JPG}
+\caption{Bronchial tree.}
+\end{figure}
+
+% First, the interaction between the viewer and the breathing noises was much more intuitive.
+In prior user interviews we found that having the intensity of the breathing incerase in the presence of the viewer was more intuitive than the intensity of breath decreasing.
+This proved to be true as the change in intensity of breathing piqued viewers interest, causing them to approach the work.
+When they were vey close, the breathing reached its highest level of intensity and many viewers noted that they felt somewhat uncomfortable to be in such close proximity to the piece.
+In comparison, some users mentioned that at far away distances -- producing slow and relaxed breathing noises -- they found their own breath falling into a relaxing rythym and were enjoyed the meditative quatily produced.
+Overall, the auditory interaction consistently caused viewers to consider their own breath.
+
+\section{Discussion}
+
+Initial research into the effects of disembodied breath revealed that the intensity and volume of the breathing noise has a large impact on the comfort level of the viewer.
+It was clear from the repetition of these trials in different settings that volume played major a role in the level of discomfort.
+Loud breathing noises inherently felt closer and more intimate than quieter noises.
+Cadence and pitch of breathing noises also appeared to have a lot of bearing on the emotion of the observer.
+A fast cadence of slightly higher pitch at higher volume was more noticeable and produced a higher degree of discomfort than a slower, softer breathing noise.
+
+In trials which involved playing noises of ``sick'' breathing noises, viewer response was much more intense.
+Results of these auditory trials were interesting because they illustrated just how sensitive the group had become to the noise of respiratory distress.
+While direct comparison is not possible, I wonder if these experiments would have had the same effect before the COVID-19 pandemic.
+
+These results were very helpful as I began to develop the digital breath generator.
+The ability to dynamically change the intensity and volume of breathing noises based on user proximity created a much more intuitive interaction between the viewer and the breathing noise.
+In prior user interviews I found that having the intensity of the breathing increase in the presence of the viewer was more natural than the intensity of breath decreasing.
+This proved to be true as the change in intensity of breathing piqued viewers interest, causing them to approach the work.
+When they were very close, the breathing reached its highest level of intensity and many viewers noted that they felt somewhat uncomfortable to be in such close proximity to the piece.
+In comparison, some viewers mentioned that at far away distances -- producing slow and relaxed breathing noises -- they found their own breath falling into a relaxing rhythm and were enjoyed the meditative quality produced.
+Overall, the auditory interaction consistently caused viewers to consider their own breath.
+
+While the variation in breath noises that could be produced using the initial trachea and diaphragm mechanism was very satisfactory, the look and feel of the mechanism had little relationship to the human body.
+Viewers noted that it was hard to relate the noised produced to the object itself.
+By making the movement of the diaphragm visible and creating an internal structure reminiscent of the bronchial tree I was able to draw viewers attention to the intended functionality of the work without the additional context of movement or sound.
+Upon seeing the bronchial structure many viewers immediately concluded that the work had to do with the lungs.
+This static visual cue really helped to enforce the intent of the interaction.
+
+During user testing when the lever mechanism was activated viewers were extremely curious about the movement of the diaphragm.
+This simple motion was the most intuitive and seemed to produce the most natural reference to breath.
+Based on the sound emanating from the bronchial tree and trachea it was very clear that the motion of diaphragm was causing air to be forced out of the vessel.
+The look of the movement of the lever head through the latex diaphragm seemed both to attract and repel viewers.
+This slightly morbid curiosity was extremely effective at holding viewer's attention on the work.
+
+%The bronchial tree added an immediate visual indicator of the work's functionality while the addition of the clear chamber allowed the viewer to guess at the diaphragm's functionality.
 \section{Further Work}
 
-well, there's always more to do...
+% emotion of breath
+% mechanical complexity
+% context
 
-the boris vian reference
+While the combination of mechanical breath and user interaction are core components of this exploration, they are not strictly necessary to continue exploring the emotion which can be conveyed through breath.
+The auditory trials in this investigation only scratched the surface of the nuanced emotion that breath can signify.
+I am curious to use this work as a platform to explore the narrative qualities of breath through viewer interaction.
+What kind of dialog can the work spark between itself and the viewer or between multiple viewers?
+What types of stories can the viewer intuit between themself and the work?
+These questions ultimately ask what range of emotion can be conveyed through the act of breathing.
+
+Expanding on the mechanical techniques developed here, I hope to create a fully automated, mechanical version of this work.
+Combining the interactive qualities of the digital breath generator with the visual qualities of the mechanical trials will ultimately tie the core interaction and the kinetic intent of this work together.
+This investigation intended to pinpoint the relationship between what is natural and what is alive through the medium of movement.
+User testing revealed that the kinetic potential of this work was the most exciting part.
+However, this work is not yet living since it does not initiate the visible act of breathing.
+
+By replacing the lever mechanism with a crank shaft -- a technique for translating circular motion into linear motion -- I hope to have fine grained control over the movement of the diaphragm and thereby the actuation of air through the work.
+This will then allow for further development and testing of the mechanical trachea.
+The modulation of breath does not end with the trachea however, and I hope to continue exploring the ways in which this sound can be manipulated.
+Perhaps the development of a mechanical pharynx, or larynx would follow.
+
+Finally, I am curious to investigate how the context of this work can be expanded beyond a traditional gallery space.
+One of the original questions this work intended to address, albeit obliquely, was how breath can be extended as a metaphor for humanity's relationship with the natural world.
+I believe it will be difficult to answer this question more specifically within the constraints of a gallery.
+Putting this piece in direct conversation with the natural world would expand the relationship of its breath beyond the viewer to the environment it inhabits.
 
@@
 The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood.
-The goal of this work was to isolate the mechanical components which form the noise created by breathing.
+The mechanical goal of this work was to isolate the components of this complex system which form the noise created by breathing.
+These investigations revealed that breath is a sensitive topic and can be interpreted as an indicator of many things.
+In addition my intent to use breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health.
+The sound of breath can be used to great effect, however, these investigations highlighted that the movement of breath is a much more intuitive signifier of life.
 
-bring this back to the question of the relationship between what is natural and what is alive
+Each human breath on this planet has moved us one step further into the Anthropocene and closer, some would argue to the end of nature.
+To conceptualize the impact of human activity on this planet is practically impossible, yet we find ourselves at a juncture where this definition is at the heart of nearly every social and political issue.
+Like James Prosek, we must all be wondering ``what can art provide in this time when our earth is indeed looking vulnerable to rapid and regrettable change induced by human influence?''\cite{prosek} If Art is a transcendent mechanism for collapsing the scales of Time and space we need it now more than ever to make sense of this recursively enormous question.
 
-here is where I can cite some of the exciting reading I've done and talk about the big distinction of artificial and natural
+% Humans have created many mechanisms to mimic various aspects of the respiratory system for purposes ranging from medical devices, underwater exploration, to speech synthesis and more.
+Most of these devices are highly specialized and seek to mimic particular aspects of respiration such as oxygen delivery, or vocal modulation.
+These works rarely mimic the actual mechanics of the components of the human body responsible for respiration favoring more abstracted and efficient mechanical systems to achieve their goal.
+Using the freedom afforded by kinetic sculpture, this work attempts to blend a human-biomimetic approach -- highlighting the mechanical simplicity of the human body -- with a highly technical approach which allows for precise and complex control over the resulting breath.
+% clearly breath is a sensitive topic and can convey many things among them emotion, health etc
+% this work confirmed that the movement of breath is a visual indicator of what it means to be alive
+% as carl zimmer notes however, life cannot be categorized by one property, it perpetually depends on ...a system
+% Using the human act of breathing as a lens, this work attempts to lay a foundation for inquiry into the relationship between what is considered to be natural and what is or has been alive.
+While the focus of this work is on a specifically human form of breath, respiration in broader context is omnipresent in living organisms and is therefore an apt biological process with which to address this question.
+This machine may breath in a natural way, but its appearance clearly identifies it as something which is purely artificial.
+
+% Mark Dion
+% Your question answers itself - the job of the artist-naturalist is precisely to interrogate the question of nature when it is subject to rapid change and its very definition is evolving in unprecedented ways.
+
+% bring this back to the question of the relationship between what is natural and what is alive
+% here is where I can cite some of the exciting reading I've done and talk about the big distinction of artificial and natural
+
+% CONCLUSION:
+% With each human breath on this planet we move deeper into the anthropocene.
+% Respiration in breathing organisms is a visibly mechanical process. 
+%
+
+% The root of my curiosity in addressing the boundary between what is natural and artificial is a desire to understand how creative practitioners can use the uncertainty surrouding thie distinction to generate reflection and consideration of humanity's place in the natural world.
 
 \cite{zimmer}
-\cite{haraway}
-\cite{smith}
 

### 2021-12-16 finalproject/report/ms1report.tex
small update to bib
close to done

@@
 Respiration is an indicator of life at all scales - from single celled organisms to human beings.
-Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it.
-Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
+Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it.\cite{zimmer} Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive.
 This mechanism may breathe in a natural way, but its appearance clearly identifies it as something which is purely artificial.
@@
 
-While not the most complex respiratory system on the planet, the human respiratory system is very complicated.\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
+The complexity of the human respiratory system lies not in its gross mechanics, but in the biological materials and molecular mechanics that allow for the transfer of oxygen to the blood.\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems.
 Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.
@@
 \includegraphics[width=0.5\textwidth]{images/buckettest.png}
-\caption{Passesby observing a breathing bucket.}
+\caption{Passerby observing a breathing bucket.}
 \end{figure}
@@
 
-The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood.
-The mechanical goal of this work was to isolate the components of this complex system which form the noise created by breathing.
+The primary objective of this work was to create a kinetic sculpture which appeared to be alive and breathing, thereby asking what it means for an non-living, artifical construct to breathe in a natural, life-like way.
+This objective was approached trough a combination of auditory and mechanical explorations.
 These investigations revealed that breath is a sensitive topic and can be interpreted as an indicator of many things.
-In addition my intent to use breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health.
-The sound of breath can be used to great effect, however, these investigations highlighted that the movement of breath is a much more intuitive signifier of life.
+In addition to my intent of using breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health.
+The ongoing COVID-19 pandemic has certainly made us more attuned to the subtle changes in breath that signal respiratory distress.
+
+The sound of breath can be used to great effect, however, these investigations revealed that the movement of breath is a much more intuitive signifier of life.
+The mechanical goal of this work was to isolate the components of the respiratory system which form the noise created by breathing and highlight their movement.
+The intensity of viewer's reactions to the mechanical system developed confirmed that the movement is just as powerful as sound in indicating life.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{ms1report}

### 2021-12-17 finalproject/report/ms1report.tex
phew

@@
 Movement is a key indicator of life and the ability to distinguish between the living and non-living is a deeply rooted capacity of the human brain.
-Psychologists and Neurologists have long been aware that ``the distinction between living and non-living things forms a natural place,'' to divide the functionalities and capacities of the mind.\cite{narine} Kinetic sculpture layers these two distinctions -- between the natural and the artificial and the living and non-living -- in an open-ended way which scientific disciplines are rarely allowed.
+Psychologists and Neurologists have long been aware that ``the distinction between living and non-living things forms a natural place,'' to divide the functionalities and capacities of the mind.\cite{narine} Kinetic sculpture layers these two distinctions -- between the natural and the artificial and the living and non-living.
 
@@
 
-Finally, it would be an oversight not to mention the hundreds of engineers and makers who devoted much time and though to the development of OpenSource emergency ventilators during the beginning and worst months of the pandemic.\cite{read} Very few of these devices were ever used on patients, but the collective thought and conversation surrounding the mechanics of breath cannot be discounted as useless, if only as a poignant reminder of the body's mystery and vulnerability.
+Finally, it would be an oversight not to mention the hundreds of engineers and makers who devoted much time and thought to the development of OpenSource emergency ventilators during the beginning and worst months of the pandemic.\cite{read} Very few of these devices were ever used on patients, but the collective thought and conversation surrounding the mechanics of breath cannot be discounted as useless, if only as a poignant reminder of the body's mystery and vulnerability.
 
@@
 This is done by the lungs and is only necessary for the transfer of oxygen to the blood.
-While the lungs are arguably the most critical part of the respiratory system, they are the most difficult to simulate with accuracy and have relatively little bearing on the sound of breath in this context.
+While the lungs are arguably the most critical part of the respiratory system, however they are the most difficult to simulate with accuracy.
+Interestingly they have relatively little bearing on the sound of breath in this context.
 
@@
 Modulation is the way the sound of breath is shaped as it leaves the body.
-This is achieved by the bronchial tree, trachea, larynx, pharynx, and mouth.\cite{sarkar} After actuation has been achieved, simulation of this system is the most relevant and interesting part of this project.
+This is achieved by the bronchial tree, trachea, larynx, pharynx, and mouth.\cite{sarkar} Changes in the shape of these structures, particularly the trachea and the mouth, change the way breath sounds as it leaves the body.
+The main focus of this project was to reproduce the sound of human breath.
+After actuation has been achieved, modulation is therefore the most interetsing and relevant part of the respiratory system to investigate.
 
@@
 
-Focusing on the two main areas of inquiry, the making and testing has been divided into been auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism.
+My two main areas of inquiry, were auditory prototypes -- exploring the effect of disembodied breath noises -- as well as mechanical prototypes -- exploring ways to achieve a physiologically accurate breathing mechanism.
 In addition to creating a visual tie between the biological nature of the respiratory system, I chose to closely mimic the diaphragm and trachea to prevent the noise of the mechanical components from overwhelming the noise of the breath produced.
@@
 In one bucket, a hole was drilled and a large diameter latex tube was inserted.
-This provided a way to direct the sound emanating from the speaker and added obfuscated the function of the assembly.
+This provided a way to direct the sound emanating from the speaker and obfuscated the function of the assembly.
 The assembly was then placed near a crosswalk with a high level of foot traffic around 8 PM ET on the day of Halloween.
@@
 Using the systems breakdown discussed above, I attempted to create mechanisms for both modulation and actuation of breath.
-I first approached the diaphragm since actuation was a necessary step to fully explore modulation.
-
-Using the five gallon buckets earlier, I cut out the center of the lid of one bucket and used it to secure a large piece of 25 mil natural latex over the opening.
+I first approached the diaphragm because actuation occurs before modulation.
+Using the five gallon buckets, I cut out the center of the lid of one bucket and used it to secure a large piece of 25 mil natural latex over the opening.
 I then inserted the same large piece of latex tubing into a hole at the other end of the bucket.
-By depressing the latex air was forced out of the tube.
+By depressing the latex, air was forced out of the tube.
 Pinching the tube at different heights and changing the cadence of the diaphragm movement created a pretty wide range of breath-like noises.
@@
 
-After measuring the force required to depress the lever fully -- 100 N -- I determined that creating an electronic system to activate the diaphragm would be costly and complicated given the scope of time.
+After measuring the force required to depress the lever fully -- 100 Newtons -- I determined that creating an electronic system to activate the diaphragm would be costly and complicated given the scope of time.
 A few attempts at creating mechanical advantage for the lever were attempted, but none had successful results.
@@
 I embedded four sections of this tubing between two sheets of fourteen mil natural latex.
-I then bent this into a curved tube-like shape and enclosed the open end of the C FIG.
+I then bent this into a curved tube-like shape and enclosed the open end of the C.
 Once this basic shape was complete I threaded each piece of tubing with monofilament and anchored the assembly between two dowels.
@@
 
-While this result was not integrated into the modulation prototype, it was by far the most naturalistic movement produced during the investigation.
+While this trachea meachanism was not integrated into the modulation prototype, it was by far the most naturalistic movement produced during the investigation.
 In comparison to the latex tubing used to mimic the trachea earlier, this mechanical latex assembly has much thinner walls.
@@
 I first created a digital breath generator in order to create a more nuanced interaction between the viewer and the breathing noises.
-This was done using P5.js and allowed me to change the cadence, pitch, and volume of the breathing noise.
+This was done using the javascript library P5.js and allowed me to change the cadence, pitch, and volume of the breathing noise.
 A simple infra-red distance sensor was then connected to the generator.
@@
 The mechanical form of the work stayed the same in this trial, however I removed the five gallon bucket in favor of a clear vessel which allowed the viewer to see the diaphragm moving.
-I also added an abstracted bronchial tree inside the vessel to suggest lungs.
-In combination with the added electronics this had a much more polished effect.
+I also added an abstract bronchial tree inside the vessel to suggest lungs.
+In combination with the added electronics this had a more polished effect.
 
@@
 
-Initial research into the effects of disembodied breath revealed that the intensity and volume of the breathing noise has a large impact on the comfort level of the viewer.
-It was clear from the repetition of these trials in different settings that volume played major a role in the level of discomfort.
+My initial research into the effects of disembodied breath revealed that the intensity and volume of the breathing noise has a large impact on the comfort level of the viewer.
+It was clear from the repetition of these trials in different settings that volume played a major role in the level of discomfort.
 Loud breathing noises inherently felt closer and more intimate than quieter noises.
-Cadence and pitch of breathing noises also appeared to have a lot of bearing on the emotion of the observer.
+Cadence and pitch of breathing noises also impacted the emotion of the observer.
 A fast cadence of slightly higher pitch at higher volume was more noticeable and produced a higher degree of discomfort than a slower, softer breathing noise.
 
-In trials which involved playing noises of ``sick'' breathing noises, viewer response was much more intense.
+In trials which involved playing ``sick'' breathing, viewer response was much more intense.
 Results of these auditory trials were interesting because they illustrated just how sensitive the group had become to the noise of respiratory distress.
@@
 
-These results were very helpful as I began to develop the digital breath generator.
+Understanding the different ways viewers reacted to breathing noises was very helpful as I began to develop the digital breath generator.
 The ability to dynamically change the intensity and volume of breathing noises based on user proximity created a much more intuitive interaction between the viewer and the breathing noise.
-In prior user interviews I found that having the intensity of the breathing increase in the presence of the viewer was more natural than the intensity of breath decreasing.
-This proved to be true as the change in intensity of breathing piqued viewers interest, causing them to approach the work.
+In user interviews I found that it was more impactful to have the breath noise increas in intensity as the user approached -- the change in intensity of breathing piqued viewers interest, causing them to approach the work.
 When they were very close, the breathing reached its highest level of intensity and many viewers noted that they felt somewhat uncomfortable to be in such close proximity to the piece.
@@
 
-By replacing the lever mechanism with a crank shaft -- a technique for translating circular motion into linear motion -- I hope to have fine grained control over the movement of the diaphragm and thereby the actuation of air through the work.
+By replacing the lever mechanism with a crank shaft -- a technique for translating circular motion into linear motion -- I hope to have fine control over the movement of the diaphragm and thereby the actuation of air through the work.
 This will then allow for further development and testing of the mechanical trachea.
@@
 
-The primary objective of this work was to create a kinetic sculpture which appeared to be alive and breathing, thereby asking what it means for an non-living, artifical construct to breathe in a natural, life-like way.
-This objective was approached trough a combination of auditory and mechanical explorations.
-These investigations revealed that breath is a sensitive topic and can be interpreted as an indicator of many things.
-In addition to my intent of using breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health.
-The ongoing COVID-19 pandemic has certainly made us more attuned to the subtle changes in breath that signal respiratory distress.
+The primary objective of this work was to create a biomimetic kinetic sculpture which appeared to be alive and breathing, thereby asking what it means for an non-living, artifical construct to breathe in a natural, life-like way.
+This objective was approached through a combination of auditory and mechanical explorations.
+These investigations revealed that both the sound and movement of breath can illicit strong emotions.
 
-The sound of breath can be used to great effect, however, these investigations revealed that the movement of breath is a much more intuitive signifier of life.
-The mechanical goal of this work was to isolate the components of the respiratory system which form the noise created by breathing and highlight their movement.
-The intensity of viewer's reactions to the mechanical system developed confirmed that the movement is just as powerful as sound in indicating life.
+As an audible indicator of life, I found that breath can be manipulated to indicate anxiety, calm, or signal respiratory distress.
+The ongoing COVID-19 pandemic has undoubtably made us more attuned to the subtle changes in breath that signal sickness or health, as was clear from my explorations of breathing noises that included coughing an wheezing.
+Yet, only a small fraction of the range of emotion that can be conveyed by breath was explored.
+I hope to extend this work to explore the narrative interactions that can be created using breath.
 
-Each human breath on this planet has moved us one step further into the Anthropocene and closer, some would argue to the end of nature.
-To conceptualize the impact of human activity on this planet is practically impossible, yet we find ourselves at a juncture where this definition is at the heart of nearly every social and political issue.
+Moving beyond the sound of breath, this work revealed that the movement of breath is also an intuitive signifier of life.
+The mechanical goal was to isolate the components of the respiratory system which form the noise created by breathing and highlight their movement.
+The intensity of viewers' reactions to the mechanical system developed confirmed that movement is just as powerful as, if not more powerful than, sound in indicating that something is alive.
+Refining the overarching context and quality of the movement will create the inherently kinetic connection between what is natural and what is alive.
+The medium of kinetic sculpture is uniquely situated to draw this connection, narrowing our focus from the incredible biological complexity of the human body, we can view the natural qualities of breath which indicate life through the most artificial and contrived of lenses.
+
+While the simple and natural act of human breath may be just that, the external acts of human existence which follow breath quickly leave behind the common ground of what is natural.
+Each human breath on this planet has moved us one step further into the Anthropocene and closer, some would argue to the end of nature.\cite{smith} To conceptualize the impact of human activity on this planet is practically impossible, yet we find ourselves at a juncture where this definition is at the heart of nearly every social and political issue.
 Like James Prosek, we must all be wondering ``what can art provide in this time when our earth is indeed looking vulnerable to rapid and regrettable change induced by human influence?''\cite{prosek} If Art is a transcendent mechanism for collapsing the scales of Time and space we need it now more than ever to make sense of this recursively enormous question.
+
+% In addition to my intent of using breath as an indicator of life to explore the boundary between the natural and artificial...
+
+% The ongoing COVID-19 pandemic has certainly made us more attuned to the subtle changes in breath that signal respiratory distress.
 

### 2022-02-06 f21/ms1/finalproject/7x7/7x7.tex
rectructureed things

--- 2021-11-04 finalproject/7x7/7x7.tex
@@
 %----------------------------------------------
-\newpage
+\clearpage
+\subsection*{five (11/03/2021)}
+\textbf{name:} brainmelting openscad brachea exercise
+
+\textbf{soundtrack:} Palomino by Trampled by Turtles then Norman Fucking Rockwell by Lana Del Rey
+
+The intent of this prototype was to model a breachea in 3D for printing.
+I did a lot of thinking at work today and was not prepared to think more.
+Luckily, David, my brain multiplier was available.
+This is as far as I got.
+\begin{figure}[h]
+\centering
+  \includegraphics[width=\textwidth]{images/53.png}
+  \caption{the beginning of a 3d model}
+\end{figure}
+
+\clearpage
 \subsection*{five (11/03/2021)}

### 2022-02-06 f21/ms1/instructionsetsforstrangers/instructionsetsforstrangers.tex
rectructureed things

--- 2021-09-30 instructionsetsforstrangers/instructionsetsforstrangers.tex
@@
 \item
-  make the hand and activation mechnaism more visible at night
+  make the hand and activation mechanism more visible at night
 \item
-  change the activation mechnaism so that it is easier to understand how to activate the hand
+  change the activation mechanism so that it is easier to understand how to activate the hand
 \end{itemize}
@@
 
-Next we found some $1/8$ inch and $3/4$ inch plywood and created the base of the new mechanishm.
+Next we found some $1/8$ inch and $3/4$ inch plywood and created the base of the new mechanism.
 At this point we also determined the best way to make the handle to activate the hand waving easier to use.
@@
 This was a simple lever that sticks out of one side of the mechanism pulled one side of the hand waving lever down when activated.
-\href{https://drive.google.com/file/d/1A4jMRefTUcI-PZPhbS6XSqFU6V4kzS8j/view?usp=sharing}{Link to video of lever action.}
+\href{https://drive.google.com/file/d/1A4jMRefTUcI-PZPhbS6XSqFU6V4kzS8j/view?usp=sharing}{Link to video of lever action.}  \href{https://drive.google.com/file/d/1zZMAdPojSRc_iC_QfIq4iEV6bjWMPP37/view?usp=sharing}{Link to video of the final lever mechanism.}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth, angle=270]{"images/III/leverwithbase.jpg"}
+\includegraphics[width=0.6\textwidth]{"images/III/leverswithbase.png"}
 \caption{Lever action.}
+\end{figure}
+
+Once the mechanism and base were dialed in, we painted the entire assembly using eye catching colors and a final layer of sparkly spray paint.
+The hope was that this would catch the light at night and make the hands more visible.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.5\textwidth]{"images/III/painting.JPG"}
+\includegraphics[width=0.5\textwidth]{"images/III/elwire.JPG"}
+\caption{Spray painted parts and installed EL wire.}
+\end{figure}
+
+We then added the electroluminescent wire around the perimeter of the large, waving hands and along the length of each lever to help attract more attention.
+Finally, after securing all the part in place, we were ready to install.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.5\textwidth]{"images/III/readytogo.JPG"}
+\caption{Wave machine ready to install.}
 \end{figure}
@@
 
-takeaways
+Our installation too place in two parts.
+Once on the evening of September 17th and then again during later afternoon on September 26th.
+The first installation seems to generate more interest as it was later at night and the hands were more visible.
+During the second installation interest seems to increase as it became darker, however we were not able to stay and watch any later.
 
-no one wants to push hard on the hands
+Once we arrived at the first installation site, around 9PM, our plans had to be modified slightly.
+There was a large construction pylon on one side of the intended intersection making the space to walk in the crosswalk quite narrow.
+We were concerned that adding the wave machine to this side of the intersection would interfere too much with accessibility to the crosswalk, so we chose the perpendicular crosswalk at the intersection instead.
+This was an unfortunate change as the perpendicular intersection had a large pile of bagged trash next to it which decreased the amount of space around the hand installed there.
+Nevertheless, there was considerable interest almost as soon as we put up the hands.
 
-people who were more intoxicaed we re less gentle
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.45\textwidth]{"images/III/purplehandinstalled.JPG"}
+\includegraphics[width=0.45\textwidth]{"images/III/greenhandinstalled.JPG"}
+  \caption{Hands installed on adjacent street corners around 9 PM (notice trash below green hand).}
+\end{figure}
+
+We chose to mount them high enough that they wouldn't be run into very often, but low enough that they were still visible to people on the street.
+During the second installation we tried lowering the hands on the street pole.
+Many more people ran into the lever and, while they were more visible, it wasn't entirely clear if this made it easier or harder to use them.
+
+Overall, the hands generated a lot of interest \href{https://drive.google.com/file/d/1pUGkjezofgZf2cuKnzpuoyqFLIYEEZ25/view?usp=sharing}{(see video of some "best of" interactions)}.
+The most challenging aspect of the installation process was waiting around long enough to observe good interactions.
+We would like to try installing the hands over a longer period of time and using a webcam to record all interactions.
+
+One thing that surprised us was how delicate people on the street treated the levers of the machine.
+Many people -- particularly during the second installation -- would come up to the hands and push lightly on the levers, not quite making them wave.
+For subsequent installations we plan to tune the mechanism so that it requires less force to achieve an exciting result.
+
+In conclusion, our interaction was quite effective!
+During the second installation we actually had a group of people wave at each other from across the street -- our intended result.
+By fine tuning the hands themselves, as well as our observation techniques, I think we can achieve more success in future installations.
 
